% $Id: Field_syn.tex,v 1.1 2003/03/10 21:54:20 cdeluca Exp $

%\section{Synopsis}

Field and Bundle objects provide 
data access and data handling methods
in the Earth System Modeling Framework (ESMF).
They are the highest level objects in the
Gridded Component object hierarchy.
Their methods are called both from user written
code as well as from other objects internal to the framework.
They are designed to meet the requirements as specified in the 
ESMF Requirements document~\ref{esmf_req<<fixme>>}.

A {\bf Field} 
object encapsulates computational and/or observational data together 
with the underlying grid on which it is located.  
A Field object provides the
application with both an overall logical view of the entire
dataset, as well as access to processor-local data after decomposition.
It provides methods for configuration, initialization, setting and
retrieving data values, data I/O, intercomponent data 
exchange, data regridding, and manipulation of attributes.

Groups of fields on the same underlying physical grid 
can be collected into a single
object called a {\bf Bundle}.  
A Bundle provides two major functions: it allows groups of Fields
to be manipulated using a single identifier, for example during
export or import of data between components; and it allows
data from multiple fields to be packed together in memory 
for higher locality of reference and ease in subsetting operations.
A Bundle object contains methods
for setting and retrieving constituent fields, import and export
data exchange during intercomponent communication,
regridding, data I/O, and reordering of data in memory.



