% $Id$

% Earth System Modeling Framework
% Copyright 2002-2018, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.

\subsubsection{ESMF\_LOGERR}
\label{const:logerr}

The valid values are:
\begin{description}
	\item [ESMF\_LOGERR\_PASSTHRU]
		A named character constant, with a predefined generic error message, 
		that can be used for the {\tt msg} argument in any {\tt ESMF\_Log} 
		routine.  The message indicated 
		by this named constant is ``{\it Passing error in return code}."
\end{description}

\subsubsection{ESMF\_LOGKIND}
\label{const:logkindflag}

{\sf DESCRIPTION:\\}
Specifies a single log file, multiple log files (one per PET), or no log files.

The type of this flag is:

{\tt type(ESMF\_LogKind\_Flag)}

The valid values are:
\begin{description}
   \item [ESMF\_LOGKIND\_SINGLE] 
         Use a single log file, combining messages from all of the PETs.  Not supported on some platforms.
   \item [ESMF\_LOGKIND\_MULTI]
         Use multiple log files --- one per PET.
   \item [ESMF\_LOGKIND\_MULTI\_ON\_ERROR]
         Use multiple log files --- one per PET.  A log file is only opened when a message
         of type {\tt ESMF\_LOGMSG\_ERROR} is encountered.
   \item [ESMF\_LOGKIND\_NONE]
         Do not issue messages to a log file.
\end{description}

\subsubsection{ESMF\_LOGMSG}
\label{const:logmsgflag}

{\sf DESCRIPTION:\\}
\begin{sloppypar}
Specifies a message level
\end{sloppypar}

The type of this flag is:

{\tt type(ESMF\_LogMsg\_Flag)}

The valid values are:
\begin{description}
   \item [ESMF\_LOGMSG\_INFO] 
         Informational messages
   \item [ESMF\_LOGMSG\_WARNING]
         Warning messages
   \item [ESMF\_LOGMSG\_ERROR]
         Error messages
   \item [ESMF\_LOGMSG\_TRACE]
         Trace messages
   \item [ESMF\_LOGMSG\_JSON]
         JSON format messages
\end{description}

Valid predefined named array constant values are:

\begin{description}
	\item [ESMF\_LOGMSG\_ALL]
		All messages
	\item [ESMF\_LOGMSG\_NONE]
		No messages
	\item [ESMF\_LOGMSG\_NOTRACE]
		All messages EXCEPT trace messages
\end{description}


