% $Id: app_tracking.tex,v 1.1 2005/05/05 22:55:21 nscollins Exp $

\section{Tracking and Software Task Sequencing}

The {\it Build and Test Plan} delivered at Milestone H outlines 
a development sequence for the Deployment Team that reflects requirements 
and the milestone delivery schedule.  The {\it Plan} is available via
the {\bf Milestones \& Archives} link, under {\bf Milestone H}.  
Software tasks shall be prioritized according to this 
{\it Plan} and as a result of meetings of JMC code development groups, JMC
code change review boards where they exist, and Joint Specification Team 
telecons.  For individual JMC codes, scientific 
development, external collaborations, release schedules, the maturity
and performance of ESMF, the applicability of an ESMF functionality class, 
and other factors may affect the priority affixed to incorporating ESMF into
the production version of the code.  If adoption into the production 
version of the code is delayed we expect Deployment Team members to 
continue ESMF implementation on a development branch and fold changes 
into the production model as soon as feasible.

System tests performed by JMC application groups will confirm that
the code works as expected.  In many cases testing changes due to
ESMF incorporation will occur in conjunction with testing other modifications
made to the application.  We expect this testing to follow the conventions
and procedures native to the application group.  Where these procedures 
are not well defined, we encourage application groups to follow the test 
procedures outlined in Section~\ref{sec:testing}.

\subsection{Tools for Software Tracking}

Developers on the Deployment Team will use the software tracker 
maintained by the Core Team to identify bugs, requests for features, 
and other development issues.  The Core Team tracker is described 
in Section~\ref{sec:tracking}.  Software bugs and issues that do not relate 
to ESMF are outside of the scope of this project.




















