%===============================================================================
% CVS $Id$
%===============================================================================

\section{Collaboration Environment and Communication}

\subsection{Mailing Lists}

The main ESMF technical mailing list is:
\begin{center}
\htmladdnormallink{esmf\_jst@list.woc.noaa.gov}{mailto:esmf\_jst@list.woc.noaa.gov}
\end{center}

This list handles telecon announcements, technical questions and
comments, discussions and materials relating to design and code
reviews, planning discussions, project announcements, and any other
items that concern the JST.

The list for active ESMF developers is:
\begin{center}
\htmladdnormallink{esmf\_core@list.woc.noaa.gov}{mailto:esmf\_core@list.woc.noaa.gov}
\end{center}

This list is for people who are developing code and checking
it into the ESMF repository.  Mails here may cover coordination 
issues relating to check-ins and releases, technical material that
is not yet ready for the JST, and information specific to NCAR
and particular development platforms.

Support questions should be directed to:
\begin{center}
\htmladdnormallink{esmf\_support@list.woc.noaa.gov}{mailto:esmf\_support@list.woc.noaa.gov}
\end{center}

The {\bf Users} tab on the ESMF website describes
how to submit an effective support request and outlines the ESMF support
policy.

People who are interested in occasional high-level project updates
can join the mailing list:
\begin{center}
\htmladdnormallink{esmf\_info@list.woc.noaa.gov}{mailto:esmf\_info@list.woc.noaa.gov}
\end{center}

Subscribers to this list receive more or less quarterly newletters
describing ESMF achievements and events.

To subscribe to any of these lists on-line, see the {\bf Users} 
tab on the ESMF website.

\subsection{Meetings and Telecons}

ESMF JST telecons are held as needed, typically on a weekly basis.
Normal telecon times are 1:00pm MT Thursdays, and sometimes Tuesdays.
A calendar of telecon topics is maintained on the home page of 
the ESMF website.  These telecons are open and are announced to 
the esmf\_jst@list.woc.noaa.gov list.

The Core Team meets weekly, at 9:30am MT on Wednesdays.  The
meeting is also set up as a telecon.  Core Team meetings are open
to active developers.

The ESMF project has open community meetings on an annual basis,
usually during late spring.  

CRB meetings are held quarterly and are closed.  However, prior 
to each CRB meeting a JST telecon is devoted to collecting comments,
requirements, and priorities from JST members for consideration by
the CRB.  

\subsection{SourceForge Open Source Development Environment}
The ESMF project utilizes the SourceForge open source development
environment.  It provides a variety of tools as well as web-browsable
repositories.

\subsubsection{The Main ESMF Site and Repository}

The main ESMF SourceForge site is at:
\begin{center}
\htmladdnormallink{{\bf http://sourceforge.net/projects/esmf}}{http://sourceforge.net/projects/esmf}
\end{center}

This site is accessible from the ESMF website,
via either the {\bf Developers} link or the {\it SourceForge}
logo on the navigation bar.  It is used for
\begin{itemize}
\item hosting and browsing the ESMF source code CVS repository; 
\item maintaining task lists;
\item archiving mailing lists;
\item providing tarballs of public source code releases (releases
      are also made available on the ESMF website);
\item tracking bugs; and 
\item tracking support requests.
\end{itemize}

The source code and tools on this site are maintained by the
ESMF Core Team, and contributions and changes cannot be made to
them without coordination with the Core Team.

\begin{sloppypar}
The main ESMF CVS repository is web-accessible at:
\htmladdnormallink{{\bf http://sourceforge.net/cvs/?group\_id=38089}}
{http://sourceforge.net/cvs/?group_id=38089} 
\end{sloppypar}

\begin{sloppypar}
The SourceForge site has instructions for checking out code.
Other CVS documentation is available at:
\htmladdnormallink{{\bf http://www.gnu.org/manual/cvs/html\_node/cvs\_toc.html}}{http://www.gnu.org/manual/cvs/html_node/cvs_toc.html} 
\end{sloppypar}

All ESMF documents, source code and test and other scripts are stored
in the main repository.  

The SourceForge repository contains only the ESMF framework.  Components
that use ESMF are stored in repositories at their home institutions.

\subsubsection{The ESMF Contributions Site and Repository}
A second ESMF SourceForge site and repository is at:
\begin{center}
\htmladdnormallink{{\bf http://sourceforge.net/projects/esmfcontrib}}{http://sourceforge.net/projects/esmfcontrib}
\end{center}

Contributors in the broader community can use this site
to archive and share code related to ESMF.  Coordination is 
not required with the ESMF Core Team in order to check into the
contributions repository.






