% $Id$

%\subsubsection{Restrictions and Future Work}

\begin{enumerate}

\item {\bf Support is limited to 1D and 2D regridding.}  Regridding
support is limited to two dimensions.

\item {\bf Masks are not implemented.}  Regridding methods take
masks in their argument lists, but they currently are not used or
applied.

\item {\bf Regridding only fills computational domains.}  Currently,
regridded values are not automatically applied to halo or ghost
domains.  Users must manually call Halo after regridding in order
to do so.  This also means that Grid periodicity will not affect
regridding results without manual Halo calls.

\item {\bf Special pole treatment is not implemented.}
Conservative regridding methods do not yet have the special pole
treatment completed.  Please see the conservative regrid algorithm
description (Section~\ref{sec:ConserveRegrid}) for further details.

\item {\bf Interpolation weights are not available to users.}  
Currently, there is no method to allow users access to the interpolation
weights for any desired manipulations.

\item {\bf Spherical coordinates are not shifted.}
Regrid methods do not yet shift longitude coordinates
(or their equivalent) from the source Grid to place them on the same
360 degree (or 2*pi) range as the destination Grid.  For Grids that
cover the entire sphere and assume a periodic longitude boundary,
this may produce errors along the Grid edges even though they share
the same 360 degree range.

\item {\bf Future regrid methods.}  The following methods will
be added:
 \begin{description}
   \item [ESMF\_REGRID\_METHOD\_ADJOINT]
         Create adjoint of existing regrid
   \item [ESMF\_REGRID\_METHOD\_FILE]
         Read a regrid from a file
   \item [ESMF\_REGRID\_METHOD\_FOURIER]
         Fourier transform
   \item [ESMF\_REGRID\_METHOD\_INDEX]
         Index-space regrid (shift, stencil)
   \item [ESMF\_REGRID\_METHOD\_LEGENDRE]
         Legendre transform
   \item [ESMF\_REGRID\_METHOD\_NEARNBR]
         Nearest-neighbor dist-weighted avg
   \item [ESMF\_REGRID\_METHOD\_RASTER]
         Regrid by rasterizing domain
   \item [ESMF\_REGRID\_METHOD\_REGRIDCOPY]
         Copy existing regrid
   \item [ESMF\_REGRID\_METHOD\_SHIFT]
         Shift addresses of existing regrid
   \item [ESMF\_REGRID\_METHOD\_SPLINE]
         Cubic spline for 1-d regridding
   \item [ESMF\_REGRID\_METHOD\_USER]
         User-supplied method
  \end{description}

\end{enumerate}


