%                **** IMPORTANT NOTICE *****
% This LaTeX file has been automatically produced by ProTeX v. 1.1
% Any changes made to this file will likely be lost next time
% this file is regenerated from its source. Send questions 
% to Arlindo da Silva, dasilva@gsfc.nasa.gov
 
\setlength{\parskip}{0pt}
\setlength{\parindent}{0pt}
\setlength{\baselineskip}{11pt}
 
%--------------------- SHORT-HAND MACROS ----------------------
\def\bv{\begin{verbatim}}
\def\ev{\end{verbatim}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\bi{\begin{itemize}}
\def\ei{\end{itemize}}
\def\bn{\begin{enumerate}}
\def\en{\end{enumerate}}
\def\bd{\begin{description}}
\def\ed{\end{description}}
\def\({\left (}
\def\){\right )}
\def\[{\left [}
\def\]{\right ]}
\def\<{\left  \langle}
\def\>{\right \rangle}
\def\cI{{\cal I}}
\def\diag{\mathop{\rm diag}}
\def\tr{\mathop{\rm tr}}
%-------------------------------------------------------------

\markboth{Left}{Source File: NUOPC\_Mediator.F90,  Date: Mon Mar 28 16:35:43 PDT 2011
}

\bigskip{\sf MODULE:}
\begin{verbatim}  module NUOPC_Mediator
\end{verbatim}

\bigskip{\sf DESCRIPTION:\\}
Mediator component with a default {\em explicit} time dependency. Each time the {\tt Run} method is called, the time stamp on the imported Fields must match the current time (on both the incoming and internal clock). Before returning, the Mediator time stamps the exported Fields with the same current time, before advancing the internal clock one timeStep forward.

\bigskip{\sf SUPER:}
\begin{verbatim}  NUOPC_ModelBase
\end{verbatim}

\bigskip{\sf USE DEPENDENCIES:}
\begin{verbatim}  use ESMF
\end{verbatim}

\bigskip{\sf SETSERVICES:}
\begin{verbatim}  subroutine SetServices(mediator, rc)
    type(ESMF_GridComp)   :: mediator
    integer, intent(out)  :: rc
\end{verbatim}

\bigskip{\sf INITIALIZE:}
\begin{itemize}
\item phase 0: ({\sc Required, NUOPC Provided})
  \begin{itemize}
  \item Initialize the {\tt InitializePhaseMap} Attribute according to the NUOPC Initialize Phase Definition (IPD) version 00 (see section \ref{IPD} for a precise definition). The default implementation sets the following mapping:
    \begin{itemize}
    \item IPDv00p1 = 1: ({\sc Required, Implementor Provided})
      \begin{itemize}
      \item Advertise Fields in import and export States.
      \end{itemize}
    \item IPDv00p2 = 2: ({\sc Required, Implementor Provided})
      \begin{itemize}
      \item Realize the advertised Fields in import and export States.
      \end{itemize}  
    \item IPDv00p3 = 3: ({\sc Required, NUOPC Provided})
      \begin{itemize}
      \item Check compatibility of the Fields' Connected status.
      \end{itemize}
    \item IPDv00p4 = 4: ({\sc Required, NUOPC Provided})
      \begin{itemize}
      \item Handle Field data initialization. Time stamp the export Fields.
      \end{itemize}
    \end{itemize}  
  \end{itemize}  
\item IPDv00p3, IPDv01p4, IPDv02p4: ({\sc NUOPC Provided})
  \begin{itemize}
  \item Set the Mediator internal clock as a copy of the incoming clock. 
  \item Check compatibility, ensuring all advertised import Fields are connected.
  \end{itemize}  
\item IPDv00p4, IPDv01p5: ({\sc NUOPC Provided})
  \begin{itemize}
  \item {\it Optional specialization} to initialize export Fields: {\tt label\_DataInitialize}
  \item Time stamp Fields in import and export States for compatibility checking.
  \end{itemize}    
\item IPDv02p5: ({\sc NUOPC Provided})
  \begin{itemize}
  \item {\it Optional specialization} to initialize export Fields: {\tt label\_DataInitialize}
  \item Time stamp Fields in export State for compatibility checking.
  \item Set Component metadata used to resolve initialize data dependencies.
  \end{itemize}    
\end{itemize}

\bigskip{\sf RUN:}
\begin{itemize}
\item phase 1: ({\sc Required, NUOPC Provided})
  \begin{itemize}
  \item {\sc Specialization Required/Provided}: {\tt label\_SetRunClock} to check and set the internal Clock against the incoming Clock.
  \begin{itemize}
  \item {\tt IF} (Phase specific specialization present): Execute the phase specific specialization.
  \item {\tt ELSE}: Execute the phase independent specialization. {\sc Provided}: By default check that internal Clock and incoming Clock agree on current time and that the time step of the incoming Clock is a multiple of the internal Clock time step. Under these conditions set the internal stop time to one time step interval of the incoming Clock. Otherwise exit with error, flagging an incompatibility.
  \end{itemize}
  \item {\sc Specialization Required/Provided}: {\tt label\_CheckImport} to check Fields in the import State.
  \begin{itemize}
  \item {\tt IF} (Phase specific specialization is present): Execute the phase specific specialization.
  \item {\tt ELSE}: Execute the phase independent specialization. {\sc Provided}: By default check that all import Fields are at the current time of the internal Clock.
  \end{itemize}
  \item Time stepping loop: starting at current time, running to stop time of the internal Clock.
  \begin{itemize}
  \item Timestamp the Fields in the export State according to the current time of the internal Clock.
  \item {\sc Specialization Required}: {\tt label\_Advance} to execute mediation code.
  \item {\sc Specialization Optional}: {\tt label\_AdvanceClock} to advance the current time of the internal Clock. By default (without specialization) advance the current time of the internal Clock according to the time step of the internal Clock.
  \end{itemize}
  \item {\sc Specialization Optional/Provided}: {\tt label\_TimestampExport} to timestamp Fields in the export State.
  \begin{itemize}
  \item {\tt IF} (Phase specific specialization present): Execute the phase specific specialization.
  \item {\tt ELSE}: Execute the phase independent specialization. {\sc Provided}: Timestamp all Fields in the export State according to the current time of the internal Clock when {\em entering} the RUN method.
  \end{itemize}
  \end{itemize}    
\end{itemize}

\bigskip{\sf FINALIZE:}
\begin{itemize}
\item phase 1: ({\sc Required, NUOPC Provided})
  \begin{itemize}
  \item {\it Optional specialization} to finalize custom parts of the component: {\tt label\_Finalize}.
  \end{itemize}      
\end{itemize}

\mbox{}\hrulefill\ 

%...............................................................
