%                **** IMPORTANT NOTICE *****
% This LaTeX file has been automatically produced by ProTeX v. 1.1
% Any changes made to this file will likely be lost next time
% this file is regenerated from its source. Send questions 
% to Arlindo da Silva, dasilva@gsfc.nasa.gov
 
\setlength{\parskip}{0pt}
\setlength{\parindent}{0pt}
\setlength{\baselineskip}{11pt}
 
%--------------------- SHORT-HAND MACROS ----------------------
\def\bv{\begin{verbatim}}
\def\ev{\end{verbatim}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\bi{\begin{itemize}}
\def\ei{\end{itemize}}
\def\bn{\begin{enumerate}}
\def\en{\end{enumerate}}
\def\bd{\begin{description}}
\def\ed{\end{description}}
\def\({\left (}
\def\){\right )}
\def\[{\left [}
\def\]{\right ]}
\def\<{\left  \langle}
\def\>{\right \rangle}
\def\cI{{\cal I}}
\def\diag{\mathop{\rm diag}}
\def\tr{\mathop{\rm tr}}
%-------------------------------------------------------------

\markboth{Left}{Source File: NUOPC\_ModelExplicitBase.F90,  Date: Mon Mar 28 16:35:43 PDT 2011
}

\bigskip{\sf MODULE:}
\begin{verbatim}  module NUOPC_ModelExplicitBase
\end{verbatim}

\bigskip{\sf DESCRIPTION:\\}
Model component that is driven by explicit time stepping. Each time the {\tt Run} method is called the model integrates one timeStep forward on the provided Clock. The Clock must be advanced between {\tt Run} calls. The component's {\tt Run} method flags incompatibility if the current time of the incoming Clock does not match the current time of the model.

\bigskip{\sf SUPER:}
\begin{verbatim}  ESMF_GridComp
\end{verbatim}

\bigskip{\sf USE DEPENDENCIES:}
\begin{verbatim}  use ESMF
\end{verbatim}

\bigskip{\sf SETSERVICES:}
\begin{verbatim}  subroutine routine_SetServices(gcomp, rc)
    type(ESMF_GridComp)   :: gcomp
    integer, intent(out)  :: rc
\end{verbatim}

\bigskip{\sf INITIALIZE:}
\begin{itemize}
\item phase 0: {\sc Required:} 
  \begin{itemize}
  \item Advertise Fields in import and export States.
  \end{itemize}
\item phase 1: {\sc Required:}
  \begin{itemize}
  \item Realize advertised Fields in import and export States by providing actual Fields.
  \end{itemize}  
\item phase 2: {\sc Provided:}
  \begin{itemize}
  \item Optionally overwrite the provided NOOP with model initialization code.
  \end{itemize}  
\item phase 3. {\sc Provided:}
  \begin{itemize}
  \item Optionally overwrite the provided NOOP with model initialization code.
  \end{itemize}  
\end{itemize}

\bigskip{\sf RUN:}
\begin{itemize}
\item phase 1: {\sc Provided:}
  \begin{itemize}
  \item Check that internal clock and incoming clock agree on current time and that the time step of the incoming clock is a multiple of the internal clock time step. Under these conditions set the internal stop time to one time step interval on the incoming clock. Otherwise exit with error, flagging incompatibility.
  \item {\it Optional specialization} to check Fields in import State: {\tt label\_CheckImport}.
  \item Model time stepping loop, starting at current time, running to stop time on the internal clock using the internal Clock time step.
  \item {\it Required specialization} to advance the model each time step: {\tt label\_Advance}.
  \item {\it Optional specialization} to timestamp the Fields in the export State: {\tt label\_TimestampExport}.
  \end{itemize}    
\end{itemize}

\bigskip{\sf FINALIZE:}
\begin{itemize}
\item phase 1: {\sc Provided:}
  \begin{itemize}
  \item Optionally overwrite the provided NOOP with model finalization code.
  \end{itemize}      
\end{itemize}

\bigskip{\sf INTERNALSTATE:}
{\sc None.}
%...............................................................

