% $Id$
%

\label{field_dictionary}

The NUOPC Layer uses standard metadata on Fields to guide the decision making that is implemented in generic code. The generic {\tt NUOPC\_Connector} component, for instance, uses the {\tt StandardName} Attribute to construct a list of matching Fields between the import and export States. The NUOPC Field Dictionary provides a software implementation of a controlled vocabulary for the {\tt StandardName} Field Attribute. It also associates each registered {\tt StandardName} with {\tt CanonicalUnits}. Currently the NUOPC Layer uses the {\tt CanonicalUnits} entry to verify that Fields are provided in their canonical units. The plan is to extend support to include unit conversion in the future.

The NUOPC Layer provides a number of default entries in the Field Dictionary, as described in section~\ref{fd:default} below. A user may extend the default Field Dictionary by calling the {\tt  NUOPC\_FieldDictionaryAddEntry()} interface to add further entries.

In alternative, a new NUOPC Field Dictionary can be setup by reading its content from a custom file in \htmladdnormallink{YAML 1.2 language}{http://yaml.org/spec/1.2/spec.html}. See section \ref{fd:custom} for details.

\subsubsection{Default Field Dictionary}
\label{fd:default}

A default NUOPC Field Dictionary is preloaded by the NUOPC Layer at start-up, and consists of the entries show in the table below.
The {\tt StandardName} Attribute of these entries complies with the Climate and Forecast (CF) conventions \htmladdnormallink{guidelines}{http://cfconventions.org/Data/cf-standard-names/docs/guidelines.html}.

\vspace{4ex}
%/////////////////////////////////////////////////////////////
\begin{tabular}{|l|l|}
\hline \hline
   {\bf StandardName}
 & {\bf CanonicalUnits}
\\ \hline \hline


                             air\_pressure\_at\_sea\_level
 &                                                Pa
\\ \hline
                  magnitude\_of\_surface\_downward\_stress
 &                                                Pa
\\ \hline
                                    precipitation\_flux
 &                                        kg m-2 s-1
\\ \hline
                    sea\_surface\_height\_above\_sea\_level
 &                                                 m
\\ \hline
                                  sea\_surface\_salinity
 &                                              1e-3
\\ \hline
                               sea\_surface\_temperature
 &                                                 K
\\ \hline
                      surface\_downward\_eastward\_stress
 &                                                Pa
\\ \hline
                     surface\_downward\_heat\_flux\_in\_air
 &                                             W m-2
\\ \hline
                     surface\_downward\_northward\_stress
 &                                                Pa
\\ \hline
                           surface\_downward\_water\_flux
 &                                        kg m-2 s-1
\\ \hline
                   surface\_eastward\_sea\_water\_velocity
 &                                             m s-1
\\ \hline
                    surface\_net\_downward\_longwave\_flux
 &                                             W m-2
\\ \hline
                   surface\_net\_downward\_shortwave\_flux
 &                                             W m-2
\\ \hline
                  surface\_northward\_sea\_water\_velocity
 &                                             m s-1
\\ \hline
\hline
\end{tabular}
%/////////////////////////////////////////////////////////////

\subsubsection{Custom Field Dictionary}
\label{fd:custom}

The default NUOPC Field Dictionary can be replaced by a custom NUOPC Field Dictionary by calling the {\tt NUOPC\_FieldDictionarySetup()} method to read in a properly-formatted YAML file. This feature is intended to improve the interoperability between codes that use the NUOPC Layer, as it allows the broader scientific community to contribute to and upkeep a common NUOPC Field Dictionary file that may be shared among different Earth System Models. At this time, an initial version of the NUOPC Field Dictionary file is available through the Earth System Community Modeling Portal (ESCOMP), a GitHub repository: \htmladdnormallink{https://github.com/ESCOMP}{https://github.com/ESCOMP}.

A custom NUOPC Field Dictionary is codified in YAML as a YAML map ({\tt key:\ value}) with only one key:~{\tt field\_dictionary}. The value associated to this key is itself a YAML map that should include the mandatory keyword {\tt entries} (pointing to the complete set of dictionary entries) and it may include the optional keywords: {\tt version\_number, last\_modified, institution, contact, source}, and {\tt description}. These optional keywords are intended to hold information about the file itself and are currently ignored by the NUOPC Layer.

Entries in the NUOPC Field dictionary are organized as YAML lists of maps. List items under the {\tt entries} keyword must be indented and marked by a hyphen ({\tt -}).

A dictionary entry fully defines a Field if it includes both the {\tt standard\_name} and {\tt canonical\_units} keywords. This entry can also include a brief narrative describing the Field as the value of optional keyword {\tt description}, which is currently ignored by the NUOPC Layer.

Synonyms can be defined by adding separate entries that include both the {\tt alias} keyword, which points to a single synonym (YAML scalar, {\it e.g.\/} {\tt alias:~<name>}) or to a comma-separated list of synonyms within square brackets (YAML flow sequence, {\it e.g.\/} {\tt alias:~[ <name1>, <name2>, ... ]}), and the {\tt standard\_name} keyword to point to the original Field name to be substituted. The original Field name must be fully defined in the dictionary file. While adding one {\tt alias} keyword to a Field definition dictionary entry is allowed and will be parsed by the NUOPC Layer, it is recommended that all synonyms be included as {\it separate} entries.

A custom NUOPC Field dictionary sample file is included below.
% this is included from src/Infrastructure/IO/tests/fd.yaml

\begin{verbatim}
field_dictionary:
    version_number: 0.0.1
    last_modified:  2018-03-14T11:01:19Z
    institution:    National ESPC, CSC & MCL Working Groups
    contact:        xfd@noaa.gov

    source:         https://github.com/ESCOMP
    description:    Community-based dictionary for shared coupling fields

    entries:
      - standard_name: air_pressure
        canonical_units: Pa
        description: Air pressure
      - standard_name: air_temperature
        canonical_units: K
        description:
          Bulk temperature of the air,
          not the surface (skin) temperature
      - alias: p
        standard_name: air_pressure
      - alias: [ t, temp ]
        standard_name: air_temperature
\end{verbatim}
