\label{ConnectorCompMeta}
The Connector Component metadata is implemented through ESMF\_Info. It can be accessed
using the JSON Pointer "/NUOPC/Instance/" prefix followed by the "Attribute name"
as per the table below. E.g. "Verbosity" is accessed using {\tt key="/NUOPC/Instance/Verbosity"}.

\begin{longtable}{|p{.25\textwidth}|p{.4\textwidth}|p{.4\textwidth}|}
     \hline\hline
     {\bf Attribute name} & {\bf Definition} & {\bf Controlled vocabulary}\\
     \hline\hline
     {\tt Kind} & String value indicating component kind.& Connector\\ \hline
     {\tt Verbosity} & String value, converted into an integer, and interpreted as a bit field. The lower 16 bits (0-15) are reserved to control verbosity of the generic component implementation. Higher bits are available for user level verbosity control. \newline
                       {\bf bit 0}: Intro/Extro of methods with indentation.\newline
                       {\bf bit 1}: Intro/Extro with memory info.\newline
                       {\bf bit 2}: Intro/Extro with garbage collection info.\newline
                       {\bf bit 3}: Intro/Extro with local VM info.\newline
                       {\bf bit 4}: Intro/Extro with ImportState info.\newline
                       {\bf bit 5}: Intro/Extro with ExportState info.\newline
                       {\bf bit 8}: Log FieldTransferPolicy.\newline
                       {\bf bit 9}: Log bond level info.\newline
                       {\bf bit 10}: Log CplList construction.\newline
                       {\bf bit 11}: Log GeomObject Transfer.\newline
                       {\bf bit 12}: Log looping over all elements in CplList for RouteHandle computation, FieldSharing, and Timestamp propagation.\newline
                       {\bf bit 13}: Log Run phase with $>>>$, $<<<$, and currTime.\newline
                       {\bf bit 14}: Log info about RouteHandle execution.\newline
                       {\bf bit 15}: Log info about RouteHandle release.
                     & 0, 1, 2, ... \newline
                       "off" = 0 (default), \newline
                       "low": some verbosity, bits: 0, 13\newline
                       "high": more verbosity, bits: 0, 4, 5, 8, 9, 10, 11, 12, 13, 14, 15\newline
                       "max": all lower 16 bits\\ \hline
     {\tt Profiling} & String value, converted into an integer, and interpreted as a bit field. The lower 16 bits (0-15) are reserved to control profiling of the generic component implementation. Higher bits are available for user level profiling control. \newline
                       {\bf bit 0}: Top level profiling of {\em Initialize} phases.\newline
                       {\bf bit 1}: Specialization point profiling of {\em Initialize} phases.\newline
                       {\bf bit 2}: Additional profiling of internals of {\em Initialize} phases.\newline
                       {\bf bit 3}: Top level profiling of {\em Run} phases.\newline
                       {\bf bit 4}: Specialization point profiling of {\em Run} phases.\newline
                       {\bf bit 5}: Additional profiling of internals of {\em Run} phases.\newline
                       {\bf bit 6}: Top level profiling of {\em Finalize} phases.\newline
                       {\bf bit 7}: Specialization point profiling of {\em Finalize} phases.\newline
                       {\bf bit 8}: Additional profiling of internals of {\em Finalize} phases.\newline
                       {\bf bit 9}: Leading barrier for {\em Initialize} phases.\newline
                       {\bf bit 10}: Leading barrier for {\em Run} phases.\newline
                       {\bf bit 11}: Leading barrier for {\em Finalize} phases.\newline
                     & 0, 1, 2, ... \newline
                       "off" = 0 (default), \newline
                       "low": Top level profiling. \newline
                       "high": Top level, specialization point profiling, and additional profiling of internals. \newline
                       "max": All lower 16 bits set.\\ \hline
     {\tt Diagnostic} & String value, converted into an integer, and interpreted as a bit field. The lower 16 bits (0-15) are reserved to control diagnostic of the generic component implementation. Higher bits are available for user level diagnostic control. \newline
                       {\bf bit 0}: Dump fields of the importState on entering {\em Initialize} phases.\newline
                       {\bf bit 1}: Dump fields of the exportState on entering {\em Initialize} phases.\newline
                       {\bf bit 2}: Dump fields of the importState on exiting {\em Initialize} phases.\newline
                       {\bf bit 3}: Dump fields of the exportState on exiting {\em Initialize} phases.\newline
                       {\bf bit 4}: Dump fields of the importState on entering {\em Run} phases.\newline
                       {\bf bit 5}: Dump fields of the exportState on entering {\em Run} phases.\newline
                       {\bf bit 6}: Dump fields of the importState on exiting {\em Run} phases.\newline
                       {\bf bit 7}: Dump fields of the exportState on exiting {\em Run} phases.\newline
                       {\bf bit 8}: Dump fields of the importState on entering {\em Finalize} phases.\newline
                       {\bf bit 9}: Dump fields of the exportState on entering {\em Finalize} phases.\newline
                       {\bf bit 10}: Dump fields of the importState on exiting {\em Finalize} phases.\newline
                       {\bf bit 11}: Dump fields of the exportState on exiting {\em Finalize} phases.
                     & 0, 1, 2, ... \newline
                       "off" = 0 (default), \newline
                       "max": All lower 16 bits set.\\ \hline
     {\tt CompLabel} & String value holding the label under which the component was added to its parent driver.& {\em no restriction}\\ \hline
     {\tt InitializePhaseMap} & List of string values, mapping the logical NUOPC initialize phases, of a specific Initialize Phase Definition (IPD) version, to the actual ESMF initialize phase number under which the entry point is registered.& IPDvXXpY=Z, where XX = two-digit revision number, e.g. 01, Y = logical NUOPC phase number, Z = actual ESMF phase number, with Y, Z > 0 and Y, Z < 10 \\ \hline
     {\tt RunPhaseMap} & List of string values, mapping the logical NUOPC run phases to the actual ESMF run phase number under which the entry point is registered.& {\em label-string}=Z, where {\em label-string} can be chosen freely, and Z = actual ESMF phase number. \\ \hline
     {\tt FinalizePhaseMap} & List of string values, mapping the logical NUOPC finalize phases to the actual ESMF finalize phase number under which the entry point is registered.& {\em label-string}=Z, where {\em label-string} can be chosen freely, and Z = actual ESMF phase number. \\ \hline
     {\tt CplList} & List of StandardNames of the connected Fields. Each StandardName entry may be followed by a colon separated list of connection options. The details are discussed in section \ref{connection_options} & {\em Standard names} as per field dictionary, followed by {\em connection options} defined in section \ref{connection_options}.\\ \hline
     {\tt CplSetList} & List of coupling sets. Each coupling set is identified by a string value.& {\em no restriction}\\ \hline
     {\tt ConnectionOptions} & String value specifying the connection options to be applied to all the fields in the {\tt CplList} by default.& {\em Connection options} defined in section \ref{connection_options}.\\ \hline
     {\tt EpochEnable} & String value specifying whether EPOCH support is enabled inside the Connector. The default setting is "true". & false, true\\ \hline
     {\tt EpochEnterKeepAlloc} & String value specifying whether to keep internal allocations when entering the EPOCH for reuse. The default setting is "true". & false, true\\ \hline
     {\tt EpochExitKeepAlloc} & String value specifying whether to keep internal allocations when exiting the EPOCH for reuse. The default setting is "true". & false, true\\ \hline
     {\tt EpochThrottle} & Integer specifying the maximum number of outstanding EPOCH messages between any two PETs. The default throttle level is 10. & Any positive integer.\\ \hline
     \hline
\end{longtable}
