\label{FieldMeta}
The Field metadata is implemented as an ESMF Attribute Package:

\begin{itemize}
    \item Convention: NUOPC
    \item Purpose: Instance
    \item Includes:
    \begin{itemize}
        \item ESG General
    \end{itemize} 
    \item Description: Basic Field description with connection and time stamp metadata. 
\end{itemize}

\begin{longtable}{|p{.22\textwidth}|p{.6\textwidth}|p{.2\textwidth}|}
     \hline\hline
     {\bf Attribute name} & {\bf Definition} & {\bf Controlled vocabulary}\\
     \hline\hline
     {\tt Connected} & Connected status.& false, true\\ \hline
     {\tt TimeStamp} & Ten integer values representing ESMF Time object.& N/A\\ \hline
     {\tt ProducerConnection} & String value indicating whether the Field has been connected with a producer. & open, targeted,\newline connected\\ \hline
     {\tt ConsumerConnection} & String value indicating whether the Field has been connected with a consumer.& open, targeted,\newline connected\\ \hline
     {\tt Updated} & String value indicating updated status during initialization.& false, true\\ \hline
     {\tt TransferOfferGeomObject} & String value indicating a component's ability to transfer information about the advertised Field, including its GeomObject.& will provide,\newline can provide,\newline cannot provide\\ \hline
     {\tt TransferActionGeomObject} & String value indicating the action a component is supposed to take with respect to transferring Field information, including its GeomObject.& provide, accept\\ \hline
     {\tt SharePolicyField} & String value indicating a component's policy with respect to sharing the Field data allocation.& share,\newline not share\\ \hline
     {\tt ShareStatusField} & String value indicating the status with respect to sharing the underlying Field data allocation that was negotiated.& shared,\newline not shared\\ \hline
     {\tt SharePolicyGeomObject} & String value indicating a component's policy with respect to sharing the Grid or Mesh on which the advertised Field object is defined.& share,\newline not share\\ \hline
     {\tt ShareStatusGeomObject} & String value indicating the status with respect to sharing the underlying GeomObject that was negotiated.& shared,\newline not shared\\ \hline
     {\tt UngriddedLBound} & Integer value list. If present equals the {\tt ungriddedLBound} of the provider field during a GeomObject transfer. & {\em no restriction}\\ \hline
     {\tt UngriddedUBound} & Integer value list. If present equals the {\tt ungriddedUBound} of the provider field.during a GeomObject transfer. & {\em no restriction}\\ \hline
     {\tt GridToFieldMap} & Integer value list. If present equals the {\tt gridToFieldMap} of the provider field.during a GeomObject transfer. & {\em no restriction}\\ \hline
     {\tt ArbDimCount} & Integer value. If present equals the {\tt arbDimCount} of the provider field.during a GeomObject transfer. & {\em no restriction}\\ \hline
     {\tt MinIndex} & Integer value list. If present equals the {\tt minIndex} (of tile 1) of the provider field.during a GeomObject transfer. & {\em no restriction}\\ \hline
     {\tt MaxIndex} & Integer value list. If present equals the {\tt maxIndex} (of tile 1) of the provider field.during a GeomObject transfer. & {\em no restriction}\\ \hline
     {\tt TypeKind} & Integer value. If present equals the integer representation of {\tt typekind} of the provider field.during a GeomObject transfer. & {\em implementation dependent range}\\ \hline
     {\tt GeomLoc} & Integer value. If present equals the integer representation of {\tt staggerloc} (for Grid) or {\tt meshloc} (for Mesh) of the provider field.during a GeomObject transfer. & {\em implementation dependent range}\\ \hline
     \hline
\end{longtable}
