% $Id$
%

\label{FieldMirror}

In some cases it is useful for a NUOPC component to match the set of fields advertised by another component, e.g. in order to connect to every field. NUOPC provides the concept of {\em field mirroring} that allows automatic matching by "mirroring" the fields of another component in their import- or exportState into their own States. One purpose of this is to automatically resolve the import data dependencies of a component, by setting up a component that exactly provides all of the needed fields.

The field mirror capability is also useful with NUOPC Mediators since these components often exactly reflect, in separate States, the sets of fields of each of the connected components.  The field mirroring capability, therefore, can be used to ensure that a Mediator is always capable of accepting fields from connected components, and removes the need to specify field lists in multiple places, i.e., both within a set of Model components connected to a Mediator and within the Mediator itself.

To access the field mirror capability, a component sets the {\tt FieldTransferPolicy} attribute during {\tt label\_Advertise}. The attribute is set on the Import- and/or Export- States to trigger field mirroring for each state, respectively.  The default value of "TransferNone" indicates that no fields should be mirrored.  The other available options are "TransferAll" and "transferAllWithNamespace". Both options mirror transfer all of the fields from all of the connected States into the State that carries the attribute. The "TransferAll" option results in flat structure with all of the mirrored fields added directly to the acceptor State. A flat structure like this is typically the preferred situation for an ExportState, where the same fields might be connected to multiple consumer components. The "transferAllWithNamespace" option also mirrors all of the field from the connected State, but creates separate Namespaces for each connection, placing the associated mirrored fields into the respective nested State. A nested structure like this useful for an ImportState where connections are being made with multiple producer components. In this case the consumer component can query the "Namespace" attribute of each nested State to infer the component label of the associated producer components.

Each Connector considers the {\tt FieldTransferPolicy} attribute on both its import and export States.  Each State that has the {\tt FieldTransferPolicy} attribute set to "transferAll" or "transferAllWithNamespace", will have then fields of the respecive other State mirror transferred. If {\em both} States have the {\tt FieldTransferPolicy} attribute set to trigger the mirror transfer, then fields are mirrored in both directions (i.e. import to export and export to import).

The transfer process works as follows:  First, the {\tt TransferOfferGoemObject} attribute is reversed between the providing side and accepting side. This is because if a field from the providing component is to be mirrored and it {\em can} provide its own geometric object, then the mirrored field on the accepting side should be set to {\em accept} a geometric object.  The mirrored field is advertised in the accepting State using a call to {\tt NUOPC\_Advertise()} such that the mirrored field shares the same StandardName.

Components have the opportunity to modify or remove any of the mirrored Fields in their Import/ExportState by using the {\tt label\_ModifyAdvertised} specialization point. After this point the initialization sequence continues as usual. Since the mirrored fields have been advertised with matching StandardName attribute, the field pairing algorithm now matches them in the usual manner, thereby establishing a connection between the original and the mirrored fields.
