%                **** IMPORTANT NOTICE *****
% This LaTeX file has been automatically produced by ProTeX v. 1.1
% Any changes made to this file will likely be lost next time
% this file is regenerated from its source. Send questions 
% to Arlindo da Silva, dasilva@gsfc.nasa.gov
 
\setlength{\parskip}{0pt}
\setlength{\parindent}{0pt}
\setlength{\baselineskip}{11pt}
 
%--------------------- SHORT-HAND MACROS ----------------------
\def\bv{\begin{verbatim}}
\def\ev{\end{verbatim}}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\bi{\begin{itemize}}
\def\ei{\end{itemize}}
\def\bn{\begin{enumerate}}
\def\en{\end{enumerate}}
\def\bd{\begin{description}}
\def\ed{\end{description}}
\def\({\left (}
\def\){\right )}
\def\[{\left [}
\def\]{\right ]}
\def\<{\left  \langle}
\def\>{\right \rangle}
\def\cI{{\cal I}}
\def\diag{\mathop{\rm diag}}
\def\tr{\mathop{\rm tr}}
%-------------------------------------------------------------

\markboth{Left}{Source File: NUOPC\_ModelExplicit.F90,  Date: Mon Mar 28 16:35:43 PDT 2011
}

\bigskip{\sf MODULE:}
\begin{verbatim} module NUOPC_ModelExplicit
\end{verbatim}

\bigskip{\sf DESCRIPTION:\\}
Model component that is driven by explicit time stepping. Each time the {\tt Run} method is called the model integrates one timeStep forward on the provided Clock. The Clock must be advanced between {\tt Run} calls. The component's {\tt Run} method flags incompatibility if the current time of the incoming Clock does not match the current time of the model.

\bigskip{\sf SUPER:}
\begin{verbatim} ESMF_GridComp
\end{verbatim}

\bigskip{\sf USE DEPENDENCIES:}
\begin{verbatim} use ESMF_Mod
\end{verbatim}

\bigskip{\sf SETSERVICES:}
\begin{verbatim} subroutine SetServices(gcomp, rc)
    type(ESMF_GridComp)  :: gcomp
    integer, intent(out) :: rc
\end{verbatim} 

\bigskip{\sf INITIALIZE:}
\begin{itemize}
\item phase 0: {\sc Required:} Add advertised Fields to import and export States.
\item phase 1: {\sc Required:} Replace advertised Fields in import and export States with actual Fields. Set the model internal Clock.
\item phase 2: {\sc Provided:} Check compatibility.
\item phase 3. {\sc Provided:} Initialize export Fields and set time stamp to current time.
\end{itemize}

\bigskip{\sf RUN:}
\begin{itemize}
\item phase 1: {\sc Provided:} Time keeping. Calls into {\em required} attachable method {\tt ModelExplicit\_Advance}.
\end{itemize}

\bigskip{\sf FINALIZE:}
\begin{itemize}
\item phase 1: {\sc Provided:} Calls into {\em optional} attachable method {\tt ModelExplicit\_Finalize}.
\end{itemize}

\bigskip{\sf INTERNALSTATE:}
{\sc None.}
%...............................................................

