% $Id: DistGrid_terms.tex,v 1.2 2002/03/11 01:07:43 vbalaji Exp $

\begin{description}
\item[PElist] a list of processor IDs associated with a component. (We
  use PE as shorthand for processor, or ``processing element'').
\item[Layout] distribution of a PElist across multiple dimensions for
  the purpose of domain decomposition.
\item[Scalar array] an array of scalar data (e.g temperature) on a
  grid.
\item[Vector array] vector data (e.g velocity components) on a
  grid. These may be arrays where one of the dimensions represents the
  vector component axis, or there may be separate data arrays for each
  vector component.
\item[Grid stencil] a descriptor of relative locations of scalar and
  vector data on a structured grid. On different staggered grids,
  vector data may lie at cell faces or vertices, while scalar data may
  lie in the interior. The staggered locations are often written
  in a notation like $(i+\frac12,j+\frac12)$ to describe the offset of
  a corner with respect to the cell $(i,j)$.
\item[Data dependency] the property of the numerics that expresses the
  requirement of non-local data values in order to perform a
  computation at a data point. For instance, a forward differencing
  operation in X at $(i,j)$ has a dependency on $(i+1,j)$.
\item[Distributed grid] Given the global range of indices associated
  with a component, this entity associates index subsets with PEs, and
  describes the the topology and connectivity of the layout.
\item[Domain decomposition] the act of grid distribution: creating a
  PE layout; and associating gridpoints with PEs.
\item[Regular distribution] the distribution along any axis is
  independent of position on any other axis. An irregular distribution
  tiles the global domain with irregularly-sized domains.
\item[Global domain] the global range of indices of data points.
\item[Computational domain] the set of points whose data is
  exclusively associated with a PE, and must be updated by it.
\item[Data domain] This includes the computational domain, as well as the
  points with whom the computational points have data dependencies.
\item[Halo] the points in the allocated domain outside the
  computational domain. Halo points are associated with other PEs'
  computational domains, and the halo update operation involves
  synchronization of halo points with other PEs. The data domain
  extents typically specify the index range used to allocated data
  arrays on a PE.
\item[Transpose] Transfer of data arrays between two distributed grids
  sharing the same global domain.
\item[Global reduction] Reduction operations (sum, max, min, etc.) on
  data on a distributed grid.
\item[Partition] in a multi-threaded application, the subset of a
  computational domain that is associated with a thread.
\item[Node]

% \item [<item1>] \label<glos:item1> <Description item 1.>

% \item [<item2>] \label<glos:item2> <Description item 2.>

\end{description}



