% $Id$

% Earth System Modeling Framework
% Copyright 2002-2021, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.

\subsubsection{ESMF\_DISTGRIDMATCH}
\label{const:distgridmatch}

{\sf DESCRIPTION:\\}
Indicates the level to which two DistGrid variables match.

The type of this flag is:

{\tt type(ESMF\_DistGridMatch\_Flag)}

The valid values are:
\begin{description}
\item [ESMF\_DISTGRIDMATCH\_INVALID:] Indicates a non-valid matching level. One
  or both DistGrid objects are invalid.
\item [ESMF\_DISTGRIDMATCH\_NONE:] The lowest valid level of DistGrid matching. 
  This indicates that the DistGrid objects don't match at any of the higher
  levels.
\item [ESMF\_DISTGRIDMATCH\_INDEXSPACE:] The index space covered by the two
  DistGrid objects is identical. However, differences between the two objects
  prevents a higher matching level.
\item [ESMF\_DISTGRIDMATCH\_TOPOLOGY:] The topology (i.e. index space and 
  connections) defined by the two DistGrid objects is identical. However, 
  differences between the two objects prevents a higher matching level.
\item [ESMF\_DISTGRIDMATCH\_DECOMP:] The index space decomposition defined by
   the two DistGrid objects is identical. However, differences between the two
   objects prevents a higher matching level.
\item [ESMF\_DISTGRIDMATCH\_EXACT:] The two DistGrid objects match in all 
  aspects, including sequence indices. The only aspect that may differ between
  the two objects is their name.
\item [ESMF\_DISTGRIDMATCH\_ALIAS:] Both DistGrid variables are aliases to the
  exact same DistGrid object in memory.
\end{description}

