% $Id: Grid_rest.tex,v 1.3 2007/05/23 15:30:40 cdeluca Exp $

%\subsubsection{Restrictions and Future Work}

\begin{itemize}

\item {\bf 7D limit.}  Only grids up to 7D will be supported.

\item {\bf During the first development phase only single
tile grids are supported.}  In the near future, support
for mosaic grids will be added.  The initial implementation 
will be to create mosaics that contain tiles of the same
grid type, e.g. rectilinear.

\item {\bf Future adaptation.}  Currently Grids
are created and then remain unchanged. In the future, it would
be useful to provide support for the various forms of grid
adaptation. This would allow the grids to dynamically change
their resolution to more closely match what is needed at a particular
time and postion during a computation for front tracking or adaptive meshes.

\item {\bf Future Grid masks.}  Grid masks will be implemented.

\item {\bf Future Exchange Grids.}  The functionality for creating an 
exchange grid between two ordinary grids will be implemented
to assist with the remapping of data during a regrid operation. 

\item {\bf Future unstructured Grid.}  Currently only grids which can be constructed from a set of logically rectangular tiles are supported. In the future more general unstructured grids will be implemented.

\item {\bf Future Grid IO.} In the future it would be useful to have a grid method which can read in a grid specification and distribution from a file and construct the grid. There may need to be a set of these methods corresponding to different group's file formats.

\item {\bf Future Grid generation.} This class for now only contains
the basic functionality for operating on the grid. In the future
methods will be added to enable the automatic generation of various types of
grids. 


\end{itemize}

