% $Id$

% Earth System Modeling Framework
% Copyright 2002-2019, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.

\subsubsection{ESMF\_PIN}
\label{const:pin_flag}

{\sf DESCRIPTION:\\}
Specifies which VM resource DEs are pinned to (PETs, VASs, SSIs).

The type of this flag is:

{\tt type(ESMF\_Pin\_Flag)}

The valid values are:
\begin{description}
  \item [ESMF\_PIN\_DE\_TO\_PET] 
    Pin DEs to PETs. Only the owning PET has access to a DE.
  \item [ESMF\_PIN\_DE\_TO\_VAS]
    Pin DEs to virtual address spaces (VAS). DEs are accessible from all PETs
    within the same VAS.
  \item [ESMF\_PIN\_DE\_TO\_SSI]
    Pin DEs to single system images (SSI) - typically shared memory nodes.
    DEs are accessible from all PETs within the same SSI. The memory allocation
    between different DEs is allowed to be non-contiguous.
  \item [ESMF\_PIN\_DE\_TO\_SSI\_CONTIG]
    Same as {\tt ESMF\_PIN\_DE\_TO\_SSI}, but the shared memory allocation
    across DEs located on the same SSI must be contigous throughout.
\end{description}


\subsubsection{ESMF\_SERVICEREPLY}
\label{const:servicereply_flag}

{\sf DESCRIPTION:\\}
Reply when a PET offers to service a DE.

The type of this flag is:

{\tt type(ESMF\_ServiceReply\_Flag)}

The valid values are:
\begin{description}
  \item [ESMF\_SERVICEREPLY\_ACCEPT] 
    The service offer has been accepted. The PET is expected to service the DE.
  \item [ESMF\_SERVICEREPLY\_DENY]
    The service offer has been denied. The PET is expected to not service the
    DE.
\end{description}

