% $Id: DELayout_usage.tex,v 1.1 2004/04/28 23:10:54 cdeluca Exp $

\subsubsection{Practical aspects of using ESMF\_DELayout}
In a typical ESMF application the application writer will create one ESMF\_DELayout for each ESMF\_Grid object, provide the ESMF\_DELayout during ESMF\_Grid creation and destroy the ESMF\_DELayout object immediately there after. The ESMF\_Grid object will hold a copy of the provided ESMF\_DELayout and a reference to this copy can be obtained using ESMF\_Grid methods anytime it is required.

The ESMF\_DELayout creation methods require a reference to the ESMF\_VM object of the current component. This reference is stored in the ESMF\_DELayout object (and its copies) and will be used in methods that require access to the virtual machine.

Typically the application writer will not need to use any of the DELayout methods directly after creating and destroying. All of the functionality provided by ESMF\_DELayout will be available through higher classes such as ESMF\_Grid and ESMF\_Field. However, in order to facilitate code migration to ESMF we do provide several ESMF\_DELayout methods which may be of interest to the application writer. Besides creation and destruction methods the available user API can be divided into four groups:
\begin{itemize}
\item {\em Get} methods which allow access to internal information,
\item {\em Print} method which allows internal information to be printed (debugging),
\item {\em Data} methods to create and destroy data objects suitable for use in communication calls,
\item {\em Communication} methods --- blocking and non--blocking point--to--point and collective communications.
\end{itemize}

\subsubsection{Examples}
{\em Here I need to fill in the layout examples I have written thus far...}
