% $Id: ArrayDataMap_options.tex,v 1.1 2004/06/04 12:29:18 nscollins Exp $

\subsubsection{ESMF\_InterleaveType}
 
{\sf DESCRIPTION:\\}
Interleave types which are used when there are multiple variables or
if individual data items are vectors.

Valid values are:
\begin{description}
   \item [ESMF\_IL\_BLOCK]
         Items are listed in blocks, all items of one type followed
         by all items of the next type.
   \item [ESMF\_IL\_ITEM]
         Items are interleaved item by item.
\end{description}


\subsubsection{ESMF\_RelLoc}

{\sf DESCRIPTION:\\}
Description of how data items are located relative to an individual
cell or element in the grid. 
(See the {\tt ESMF\_Grid} documentation
for a description of 'staggering' which is a per-grid concept.)

Valid values are:
\begin{description}
  \item[ESMF\_CELL\_UNDEFINED]  Data location is undefined.
  \item[ESMF\_CELL\_CENTER] Data location is at cell center.
  \item[ESMF\_CELL\_NFACE] Data location is on north face of cell.
  \item[ESMF\_CELL\_SFACE] Data location is on south face of cell.
  \item[ESMF\_CELL\_EFACE] Data location is on east face of cell.
  \item[ESMF\_CELL\_WFACE] Data location is on west face of cell.
  \item[ESMF\_CELL\_NECORNER] Data location is at north-east corner of cell.
  \item[ESMF\_CELL\_NWCORNER] Data location is at north-west corner of cell.
  \item[ESMF\_CELL\_SECORNER] Data location is at south-east corner of cell.
  \item[ESMF\_CELL\_SWCORNER] Data location is at south-west corner of cell.
  \item[ESMF\_CELL\_TOPFACE] Data location is on top face of cell.
  \item[ESMF\_CELL\_BOTFACE] Data location is on bottom face of cell.
  \item[ESMF\_CELL\_CELL] Data location is over entire cell.
  \item[ESMF\_CELL\_VERTEX] Data location is at the vertices of the cell.
\end{description}

\subsubsection{ESMF\_IndexOrder}

{\sf DESCRIPTION:\\}
A set of predefined index orders which shortcut setting
the mapping between data and grid indices.

\begin{description}
   \item[ESMF\_INDEX\_I]  One dimensional data and grid.
   \item[ESMF\_INDEX\_IJ] Two dimensional, IJ ordering.
   \item[ESMF\_INDEX\_JI] Two dimensional, JI ordering.
   \item[ESMF\_INDEX\_IJK] Three dimensional, IJK ordering.
   \item[ESMF\_INDEX\_JIK] Three dimensional, JIK ordering.
   \item[ESMF\_INDEX\_KJI] Three dimensional, KJI ordering.
   \item[ESMF\_INDEX\_IKJ] Three dimensional, IKJ ordering.
   \item[ESMF\_INDEX\_JKI] Three dimensional, JKI ordering.
   \item[ESMF\_INDEX\_KIJ] Three dimensional, KIJ ordering.
\end{description}

