% $Id: comp_desc.tex,v 1.2 2003/01/30 23:42:35 nscollins Exp $
%
% Earth System Modeling Framework
% Copyright 2002-2003, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the GPL.

%\subsection{Description}

A {\bf Component}
is the highest level object in the ESMF object
hierarchy.  Components are a combination of user written code and
data structures and interfaces defined by the ESMF framework.
Components must provide a set of predefined subroutine interfaces
which will be called by higher level code.  They must encapsulate
their input and output data in an ESMF {\tt State} object.  They
must able to interpret an ESMF {\tt Time} object in order
to execute the requested number of model timesteps or time interval
before exchanging data with other Components.

Components are identified as being one or more of the
following subtypes: {\bf Application}, {\bf Gridded}, and {\bf Coupler}.
An Application Component contains other Components which
cooperate in a computation. An Application Component can also be itself a
Component nested in a higher level Application.  A Gridded Component
does the actual model computation.  A Coupler Component executes or
arranges the execution of the necessary data transformations between one
or more Gridded Components.

