% $Id: Component_syn.tex,v 1.2 2003/01/30 23:42:34 nscollins Exp $
%
% Earth System Modeling Framework
% Copyright 2002-2003, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the GPL.

%\section{Synopsis}

Component, State, and Transform objects are part of the Superstructure 
layer of the Earth System Modeling Framework (ESMF).  They encapsulate
the computational models and the exchange of data between them.
They are designed to meet the requirements as specified in the
ESMF Requirements document~\ref{esmf_req}.

A {\bf Component} 
is the highest level object in the ESMF object
hierarchy.  Components are a combination of user written code and
data structures and interfaces defined by the ESMF framework.
Components must provide a set of predefined subroutine interfaces
which will be called by higher level code.  They must encapsulate
their input and output data in an ESMF {\tt State} object.  They
must able to interpret an ESMF {\tt Time} object in order
to execute the requested number of model timesteps or time interval
before exchanging data with other Components.

Components are identified as being one or more of the
following subtypes: {\bf Application}, {\bf Gridded}, and {\bf Coupler}.  
An Application Component contains other Components which
cooperate in a computation. An Application Component can also be itself a
Component nested in a higher level Application.  A Gridded Component
does the actual model computation.  A Coupler Component executes or
arranges the execution of the necessary data transformations between one
or more Gridded Components.


A {\bf State} 
encapsulates the data and sufficient description information
to allow the data computed by one Gridded Component to be
used as input to another Gridded Component.  A Gridded Component has
two states, an {\bf Import State} and an {\bf Export State}.
The Import State encapsulates all data needed as input to the
next run of the Component.  The Export State encapsulates all
data produced by the Component which is required by another
Component as input.  Data inside states are one or more of
the following ESMF Objects: Bundles, Fields, or Arrays.
Components must exchange information with other Components
only by passing the data through State objects.

A Coupler Component takes a list of State objects consisting of
both Export States and Import States.
The Coupler Component is responsible for transforming data 
so the appropriate Import States can be provided the corresponding
Component or Components.
The framework supplies
methods to do basic transformations which may include regridding,
interpolation, unit conversions, accumulations, averaging,
transformations which conserve specific values.


A {\bf Transform} object encapsulates a list of operations to be
performed on a {\tt State} object.  For Components where it is not
feasible to return to the calling routine each time Coupling is
required, the Transform methods can be called from within the
Component to perform the proper data transformations.

