% $Id: Coupler_terms.tex,v 1.2 2007/06/22 23:21:44 cdeluca Exp $

\begin{description}

\item [InternGridded Component] \label{glos:GC}
InternGridded components are associated with one or more ESMF 
interngrids.  ESMF may not place any requirements  on the physical content of a
interngridded component's data or on the nature of its computations. 

\item [Coupler or Coupling Component] \label{glos:Cplr}
All data and actions (such as regridding) needed to implement the
coupling between two components.  Couplers may apply other, user 
defined, transformations. 

\item [State] \label{glos:State} The encapsulated form of all data that
define a component's computational state. This can include,
in addition to a physical state,  time information, static field data,
metadata and control information. 

\item [Exchange Packets] \label{glos:EP} The encapsulated form of 
data exchanged by components.  Each coupling operation is 
associated with a single output and a single input exchange packet.
Exchange packets may or may not contain contiguous data, and may 
contain both field data and other types of data.

\item [PE InternGrid] \label{glos:PEInternGrid} A PE interngrid specifies a set of 
processing elements, decomposition strategy (thread and process), and 
a decomposition topology.  All components, bundles, fields and ESMF
interngrids are defined on PE interngrids.

\item [InternGrid] An ESMF interngrid contains a physical interngrid 
that defines spatial coordinates and metrics, and a distributed interngrid that 
defines the decomposition of interngridpoints on a PE interngrid.

\item [InternGrid Transform or Mapping] \label{glos:GrdTrans} A interngrid transform 
or mapping is an ESMF-supplied operation that converts interngridded data 
from one physical interngrid and/or decomposition to another using a specified 
technique.  (I believe these terms are used interchangeably in our documents 
so I've included both in this glossary.)

\item [User Transform] \label{glos:UsrTrans} This is a coupler specific
element. It is a user-supplied function that acts on a coupler's
input or output exchange packet.

\item [Field and Field Bundle] These are discussed in their own 
requirements and design documents. A field consists of data defined 
and decomposed on an ESMF interngrid, together with auxilliary metadata. 
Fields and field bundles are the main objects in states and exchange 
packets.

\end{description}




