% +-======-+ 
%  Copyright (c) 2003-2007 United States Government as represented by 
%  the Admistrator of the National Aeronautics and Space Administration.  
%  All Rights Reserved.
%  
%  THIS OPEN  SOURCE  AGREEMENT  ("AGREEMENT") DEFINES  THE  RIGHTS  OF USE,
%  REPRODUCTION,  DISTRIBUTION,  MODIFICATION AND REDISTRIBUTION OF CERTAIN 
%  COMPUTER SOFTWARE ORIGINALLY RELEASED BY THE UNITED STATES GOVERNMENT AS 
%  REPRESENTED BY THE GOVERNMENT AGENCY LISTED BELOW ("GOVERNMENT AGENCY").  
%  THE UNITED STATES GOVERNMENT, AS REPRESENTED BY GOVERNMENT AGENCY, IS AN 
%  INTENDED  THIRD-PARTY  BENEFICIARY  OF  ALL  SUBSEQUENT DISTRIBUTIONS OR 
%  REDISTRIBUTIONS  OF THE  SUBJECT  SOFTWARE.  ANYONE WHO USES, REPRODUCES, 
%  DISTRIBUTES, MODIFIES  OR REDISTRIBUTES THE SUBJECT SOFTWARE, AS DEFINED 
%  HEREIN, OR ANY PART THEREOF,  IS,  BY THAT ACTION, ACCEPTING IN FULL THE 
%  RESPONSIBILITIES AND OBLIGATIONS CONTAINED IN THIS AGREEMENT.
%  
%  Government Agency: National Aeronautics and Space Administration
%  Government Agency Original Software Designation: GSC-15354-1
%  Government Agency Original Software Title:  GEOS-5 GCM Modeling Software
%  User Registration Requested.  Please Visit http://opensource.gsfc.nasa.gov
%  Government Agency Point of Contact for Original Software:  
%  			Dale Hithon, SRA Assistant, (301) 286-2691
%  
% +-======-+ 
{\tt MAPL\_HistoryGridCompMod} is an internal MAPL gridded component 
used to manange output streams from a MAPL hierarchy. It writes Fields in the
Export states of all MAPL components in a hierarchy to file collections
during the course of a run. It also has the some limited capability to interpolate
the fields horizontally and/or vertically beofore outputing them. 

It is usually one of the
two gridded components in the ``cap'' or main program of a MAPL application,
the other being the root of the MAPL hierarchy it is servicing. It is 
instanciated and all its registered methods are run automatically by 
{\tt MAPL\_Cap,} if that is used.
If writing a custom cap, {\tt MAPL\_HistoryGridCompMod}'s SetServices can be 
called anytime after ESMF is initialized.
Its Initialize method should be executed before entering the time loop, and its
Run method at the bottom of each time loop, after advancing the Clock. Finalize
simply cleans-up memory. 

The component has no true export state, since its products are diagnostic
file collections. It does have both Import and Internal states, which can be treated as
in any other MAPL component, but it generally makes no sense to checkpoint and restart
these.

The behavior of  {\tt MAPL\_HistoryGridCompMod} is controlled through its configuration,
which as in any MAPL gridded component, is open and available in the GC. It is placed
there by the cap and usually contained in a HISTORY.rc file.

{\tt MAPL\_HistoryGridCompMod} uses {\tt MAPL\_CFIO} for creating and writing its files;
it thus obeys all {\tt MAPL\_CFIO} rules. In particular, an application can write either
Grads style flat files together with the Grads .ctl file description files, or
one of two self-describing format (netcdf or HDF), which ever is linked with the 
application.

Each collection  to be produced is described in the HISTORY.rc file and can have the
following properties:

\begin{itemize}
\item Its fields may be "instantaneous" or "time-averaged", but all fields within
  a collection use the same time discretization. 
\item A beginning and an end time may be specified for each collection .
\item Collections are a set of files with a common name template. 
\item Files in a collection have a fixed number of time groups in them.
\item Data in each time group are "time-stamped"; for time-averaged data,
  the center of the averaging period is used.
\item Files in a collection can have time-templated names. The template
  values correspond to the times on the first group in the file.
\end{itemize}

The body of the HISTORY.rc file usually begins with two
character string attributes under the config labels {\tt EXPID:} and {\tt EXPDSC:}
that are identifiers for the full set of collections. These are followed
by a list of collection names under the config label {\tt COLLECTIONS:}. Note
the conventional use of colons to terminate labels in the HISTORY.rc.

The remainder of the file contains the attributes for each collection.
Attribute labels consist of the attribute name with the collection name
prepended; the two are separated by a '.'.

Attributes are listed below. A special attribute is {\tt {\em collection}.fields:}
which is the label for the list of fields that will be in the collection.
Each item (line) in the field list consists of a comma separated list
with the field's name (as it appears in
the corresponding ESMF field in the EXPORT of the component), the name of the component that
produces it, and the alias to use for it in the file. The alias may be omitted, in which case
it defaults to the true name.

Files in a collection are named using the collection name, the
template attribute described below, and the {\tt EXDID:} attribute
value. A filename extension may also be added to identify the type of
file (e.g., .nc4).
\begin{quote}
    {\tt [expid.]collection[.template][.ext]}
\end{quote}
The extension is not added automatically, it is up to the user to add the appropriate one.
If the format is CFIO or CFIOasync and the extension is absent or .nc a NETCDF4 classic file
will be produced. Is the extentions is .nc4 a NETCDF4 file will be produced.
If it is "flat", the data files have whatever extension you provide and
the ``control file'' has the .ctl extension, but with no {\tt
template}. The expid is always prepended, unless it is an empty
string.

 The following are the valid collection attributes:
 \begin{quote}
 \begin{trivlist}
 \item[\tt template]      Character string defining the time stamping template that is appended 
                          to {\tt collection} to create a particular file name. 
                          The template uses GrADS convensions. 
                          The default value depends on the {\tt duration} of the file.
 \item[\tt descr]         Character string describing the collection. Defaults to `expdsc'.
 \item[\tt format]        Character string to select file format ("CFIO", "CFIOasync", "flat").  "CFIO"
                          uses MAPL\_CFIO and produces netcdf output. "CFIOasync" uses MAPL\_CFIO but
                          delegates the actual I/O to the MAPL\_CFIOServer (see MAPL\_CFIOServer documenation for details).
                          Default = "flat".
 \item[\tt frequency]     Integer (HHHHMMSS) for the frequency of time groups in the collection.
                          Default = 060000.
 \item[\tt mode]          Character string equal to `instantaneous' or `time-averaged'.
                          Default = 'instantaneous'.
 \item[\tt acc\_interval] Integer (HHHHMMSS) for the acculation interval ($\le$ frequency)
                          for time-averaged diagnostics. Default = {\tt frequency}; ignored
                          if {\tt mode} is `instantaneous'.
 \item[\tt ref\_date]     Integer (YYYYMMDD) reference date for {\em frquency};
                          also the beginning date for
                          the collection. Default is the Start date on the Clock.
 \item[\tt ref\_time]     Integer (HHMMSS) Same a {\tt ref\_date}.
 \item[\tt end\_date]     Integer (YYYYMMDD) ending date to stop diagnostic output.
                          Default: no end
 \item[\tt end\_time]     Integer (HHMMSS) ending time to stop diagnostic output.
                          Default: no end.
 \item[\tt duration]      Integer (HHHHMMSS) for the duration of each file. 
                          Default = 00000000 (everything in one file).
 \item[\tt resolution]    Optional resolution (IM JM) for the ouput stream.
                          Transforms betwee two regulate LogRect grid in index space. 
                          Default is the native resolution.
 \item[\tt xyoffset]      Optional Flag for output grid offset when interpolating. Must be
                          between 0 and 3. (Cryptic Meaning: 0:DcPc, 1:DePc, 2:DcPe, 3:DePe).
                          Ignored when {\tt resolution} results in no interpolation (native).
                          Default: 0 (DatelinCenterPoleCenter). 
 \item[\tt levels]        Optional list of output levels (Default is all levels on Native Grid).
                          If {\tt vvars} is not specified, these are layer indeces. Otherwise
                          see {\tt vvars, vunits, vscale}.
 \item[\tt vvars]         Optional field to use as the vertical coordinate and functional form
                          of vertical interpolation. A second argument specifies 
                          the component the field comes from. 
                          Example 1: the entry 'log(PLE)','DYN' uses PLE from the
                          DYN component as the vertical coordinate and interpolates
                          to {\tt levels} linearly in its log. Example 2: 'THETA','DYN'
                          a way of producing isentropic output.
                          Only {\tt log}($\cdot$), {\tt pow}($\cdot$,{\em real number})
                          and straight linear interpolation are supported.
 \item[\tt vunit]         Character string to use for units attribute of the vertical 
                          coordinate in file. 
                          The default is the MAPL\_CFIO default. 
                          This affects only the name in the file.
                          It does not do the conversion. See {\tt vscale}
 \item[\tt vscale]        Optional Scaling to convert VVARS units to VUNIT units.
                          Default: no conversion.
 \item[\tt regrid\_exch]  Name of the exchange grid that can be used for interpolation
                          between two LogRect grids or from a tile grid to a LogRect grid.
                          Default: no exchange grid interpolation.
                          irregular grid.
 \item[\tt regrid\_name]  Name of the Log-Rect grid to interpolate to when going from a tile 
                          to Field to a gridde output. {\tt regrid\_exch} must be set, otherwise
                          it is ignored.
 \item[\tt conservative]  Set to a non-zero integer to turn on conservative regridding when going
                          from a native cube-sphere grid to lat-lon output.
                          Default: 0
 \item[\tt deflate]       Set deflate level (0-9) of NETCDF output when format is CFIO or CFIOasync.
                          Default: 0
 \item[\tt subset]        Optional subset (lonMin lonMax latMin latMax) for the output when performing
                          non-conservative cube-sphere to lat-lon regridding of the output.
 \item[\tt chunksize]     Optional user specified chunking of NETCDF output when format is CFIO or CFIOasync,
                          (Lon chunksize, Lat chunksize, Lev chunksize, Time chunksize)
\end{trivlist}
\end{quote}

 The following is a sample HISORY.rc take from the FV\_HeldSuarez test.
 \begin{verbatim}
  EXPID:  fvhs_example
  EXPDSC: fvhs_(ESMF07_EXAMPLE)_5x4_Deg

  COLLECTIONS:
        'dynamics_vars_eta'
        'dynamics_vars_p'
        ::
 
 
 
 
  dynamics_vars_eta.template:   '%y4%m2%d2_%h2%n2z',
  dynamics_vars_eta.format:     'CFIO',
  dynamics_vars_eta.frequency:  240000,
  dynamics_vars_eta.duration:   240000,
  dynamics_vars_eta.fields:     'T_EQ'     , 'HSPHYSICS'           ,
                                'U'        , 'FVDYNAMICS'          ,
                                'V'        , 'FVDYNAMICS'          ,
                                'T'        , 'FVDYNAMICS'          ,
                                'PLE'      , 'FVDYNAMICS'          ,
                        ::
 
 
 
  dynamics_vars_p.template:   '%y4%m2%d2_%h2%n2z',
  dynamics_vars_p.format:     'flat',
  dynamics_vars_p.frequency:  240000,
  dynamics_vars_p.duration:   240000,
  dynamics_vars_p.vscale:     100.0,
  dynamics_vars_p.vunit:      'hPa',
  dynamics_vars_p.vvars:      'log(PLE)' , 'FVDYNAMICS'          ,   
  dynamics_vars_p.levels:      1000 900 850 750 500 300 250 150 100 70 
                               50 30 20 10 7 5 2 1 0.7, (SHOULD BE IN ONE LINE??)
  dynamics_vars_p.fields:     'T_EQ'     , 'HSPHYSICS'           ,
                              'U'        , 'FVDYNAMICS'          ,
                              'V'        , 'FVDYNAMICS'          ,
                              'T'        , 'FVDYNAMICS'          ,
                              'PLE'      , 'FVDYNAMICS'          ,
                        ::
 \end{verbatim}

In addition to fields in the EXPORT state of components in the MAPL hierarchy,
the user may also specify new fields that will be evaluted using the MAPL
arithmetic parser. In place of the variable name when defining fields
the user enters an expression that will be evaluated by the MAPL parser (see the documentation for the parser
to learn what legal expressions are). In the expression the user can use any field
in the collection as a variable in the expression, provided it is an actual field
in a MAPL component. In other words new arithemtic fields CAN NOT be used as variables
in other expressions in the collection.
On this line the user must specify two more pieces of information. A component
name is needed just as in the real fields but this is just a placeholder and will be ignored. Finally a name for the
new field is needed. This will be the name of the variable on the resultant file produced by {\tt MAPL\_HistoryGridCompMod}. 
An example with some comments below makes this clear: 
 \begin{verbatim}
  EXPID:  fvhs_example
  EXPDSC: fvhs_(ESMF07_EXAMPLE)_5x4_Deg

  COLLECTIONS:
        'dynamics_vars_eta'
        ::

  dynamics_vars_eta.template:   '%y4%m2%d2_%h2%n2z',
  dynamics_vars_eta.format:     'CFIO',
  dynamics_vars_eta.frequency:  240000,
  dynamics_vars_eta.duration:   240000,
  dynamics_vars_eta.fields:     'T_EQ'     , 'HSPHYSICS'           ,
                                'U'        , 'FVDYNAMICS'          ,
                                'V'        , 'FVDYNAMICS'          ,
                                'sqrt(U*U+V*V)', 'FVDYNAMCIS', 'Wind_Magnitude',
                                'U*2.0e2', 'FVDYNAMICS', 'U_times_two', 
                        ::
 # T_EQ, U, and V are actual fields in the export state of MAPL components.
 # note two new fields that are functions of the real fields in the collection.
 # note that in the line 'sqrt(U*U+V*V)', 'FVDYNAMCIS', 'Wind_Magnitude',
 # the FVDYNAMICS entry is just a placeholder. It is not used.
 # it would be illegal to use Wind_Magnitude in an expression as a variable.
 \end{verbatim}
