% $Id$
%
% Earth System Modeling Framework
% Copyright 2002-2021, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.

%TODO: This file started as an exact copy of the Fortran version of this file.
%TODO: Changes were made to correctly reflect the current status of the C API.
%TODO: Eventually this file should be removed again and replaced by a single
%TODO: generic version that can be included for both Fortran and C refdocs.

%\subsection{Description}

\label{sec:State}

A State contains the data and metadata to be transferred between 
ESMF Components.  It is an important class, because it defines a 
standard for how data is represented in data transfers between Earth
science components.  The 
State construct is a rational compromise between a fully prescribed 
interface - one that would dictate what specific fields should be 
transferred between components - and an interface in which data structures
are completely ad hoc.

There are two types of States, import and export.
An import State contains data that is necessary for a Gridded Component
or Coupler Component to execute, and an export State contains the data
that a Gridded Component or Coupler Component can make available.

States can contain Arrays, ArrayBundles, Fields, FieldBundles, 
and other States. However, the current C API only provides State access to
Arrays, Fields and nested States.
States cannot directly contain native language arrays
(i.e. Fortran or C style arrays).  Objects in a State must span
the VM on which they are running.  For sequentially executing components
which run on the same set of PETs this happens by calling the object
create methods on each PET, creating the object in unison.   For
concurrently executing components which are running on subsets of PETs,
an additional method, called {\tt ESMF\_StateReconcile()}, is provided by
ESMF to broadcast information
about objects which were created in sub-components.
Currently this method is only available through the ESMF Fortran API. Hence
the Coupler Component reponsible for reconciling States from Component that 
execute on subsets of PETs must be written in Fortran.

State methods include creation and deletion, adding and retrieving 
data items, and performing queries.  


