% $Id$
%
% Earth System Modeling Framework
% Copyright 2002-2021, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.

%\subsection{Description}

\label{sec:State}

A State contains the data and metadata to be transferred between 
ESMF Components.  It is an important class, because it defines a 
standard for how data is represented in data transfers between Earth
science components.  The 
State construct is a rational compromise between a fully prescribed 
interface - one that would dictate what specific fields should be 
transferred between components - and an interface in which data structures
are completely ad hoc.

There are two types of States, import and export.
An import State contains data that is necessary for a Gridded Component
or Coupler Component to execute, and an export State contains the data
that a Gridded Component or Coupler Component can make available.

States can contain Arrays, ArrayBundles, Fields, FieldBundles, 
and other States.  They cannot directly contain native language arrays
(i.e. Fortran or C style arrays).  Objects in a State must span
the VM on which they are running.  For sequentially executing components
which run on the same set of PETs this happens by calling the object
create methods on each PET, creating the object in unison.   For
concurrently executing components which are running on subsets of PETs,
an additional method, called {\tt ESMF\_StateReconcile()}, is provided by
ESMF to broadcast information
about objects which were created in sub-components.

State methods include creation and deletion, adding and retrieving 
data items, adding and retrieving attributes, and performing queries.  


