% $Id: State_options.tex,v 1.7 2004/06/11 19:21:37 cdeluca Exp $

\subsubsection{ESMF\_StateType}

{\sf DESCRIPTION:\\}
Specifies whether an {\tt ESMF\_State} contains data to be imported
into a component or exported from a component. 

Valid values are:
\begin{description}
   \item [ESMF\_STATE\_IMPORT] 
         Contains data to be imported into a component.
   \item [ESMF\_STATE\_EXPORT]
         Contains data to be exported out of a component.
   \item [ESMF\_STATE\_INVALID]
         Does not contain valid data.
\end{description}

\subsubsection{ESMF\_StateItem}
\label{opt:stateitem}

{\sf DESCRIPTION:\\}
Specifies the type of object being added to or retrieved from an
{ESMF\_State.}

Valid values are:
\begin{description}
   \item [ESMF\_STATEITEM\_BUNDLE] 
         Refers to an {\tt ESMF\_Bundle} within an {\tt ESMF\_State}.
   \item [ESMF\_STATEITEM\_FIELD]
         Refers to an {\tt ESMF\_Field} within an {\tt ESMF\_State}.
   \item [ESMF\_STATEITEM\_ARRAY]
         Refers to an {\tt ESMF\_Array} within an {\tt ESMF\_State}.
   \item [ESMF\_STATEITEM\_STATE]
         Refers to an {\tt ESMF\_State} within an {\tt ESMF\_State}.
   \item [ESMF\_STATEITEM\_NAME]
         Refers to a data name used as a placeholder within 
         an {\tt ESMF\_State}.
   \item [ESMF\_STATEITEM\_UNKNOWN]
         Object type within an {\tt ESMF\_State} is unknown.
\end{description}

\subsubsection{ESMF\_StateDataNeeded}

{\sf DESCRIPTION:\\}
Specifies whether or not an {\tt ESMF\_State} entry is needed for a 
particular application configuration.

Valid values are:
\begin{description}
   \item [ESMF\_STATEDATA\_NEEDED] 
         Data is needed.
   \item [ESMF\_STATEDATA\_NOTNEEDED]
         Data is not needed.
\end{description}






