% $Id$

\section{Appendix A: Master List of Constants}
\label{const:cmaster}

\subsection{ESMC\_CALKIND}
This flag is documented in section \ref{const:calkindflag_c}.

\subsection{ESMC\_COORDSYS}
This flag is documented in section \ref{const:ccoordsys}.

\subsection{ESMC\_DECOMP}
\label{const:cdecompflag}
{\sf DESCRIPTION:\\}
Indicates how DistGrid elements are decomposed over DEs.

The type of this flag is:

{\tt type(ESMC\_Decomp\_Flag)}

The valid values are:
\begin{description}
\item [ESMC\_DECOMP\_BALANCED]
      Decompose elements as balanced as possible across DEs. The maximum 
      difference in number of elements per DE is 1, with the extra elements on
      the lower DEs.
\item [ESMC\_DECOMP\_CYCLIC]
      Decompose elements cyclically across DEs.
\item [ESMC\_DECOMP\_RESTFIRST]
      Divide elements over DEs. Assign the rest of this division to the first
      DE.
\item [ESMC\_DECOMP\_RESTLAST]
      Divide elements over DEs. Assign the rest of this division to the last DE.
\end{description}

{\tt type(ESMC\_ExtrapMethod\_Flag)}
\label{opt:cextrapmethod}

{\sf DESCRIPTION:\\}
Specify which extrapolation method to use on unmapped destination points after 
regridding.

The type of this flag is:

{\tt type(ESMC\_ExtrapMethod\_Flag)}

The valid values are:
\begin{description}
\item [ESMC\_EXTRAPMETHOD\_NONE]
      Indicates that no extrapolation should be done. 
\item [ESMC\_EXTRAPMETHOD\_NEAREST\_STOD]
      Inverse distance weighted average. 
      Here the value of a destination point is the weighted average of the 
      closest N source points. The weight is the reciprocal of the distance of 
      the source point from the destination point raised to a power P. All the
      weights contributing to one destination point are normalized so that they 
      sum to 1.0. The user can choose N and P when using this method, but 
      defaults are also provided.  
\item [ESMC\_EXTRAPMETHOD\_NEAREST\_IDAVG]
      Nearest source to destination. 
      Here each destination point is mapped to the closest source point. A given 
      source point may go to multiple destination points, but no destination 
      point will receive input from more than one source point. 
\end{description}

\subsection{ESMC\_FILEFORMAT}
This flag is documented in section \ref{const:cfileformat}.

\subsection{ESMC\_GRIDITEM}
This flag is documented in section \ref{const:cgriditem}.

\subsection{ESMC\_GRIDSTATUS}
This flag is documented in section \ref{const:cgridstatus}.

\subsection{ESMC\_INDEX}
\label{const:indexflag}

{\sf DESCRIPTION:\\}
Indicates whether index is local (per DE) or global (per object).

The type of this flag is:

{\tt type(ESMC\_IndexFlag)}

The valid values are:
\begin{description}
\item [ESMC\_INDEX\_DELOCAL]
      Indicates that DE-local index space starts at lower bound 1 for each DE.
\item [ESMC\_INDEX\_GLOBAL]
      Indicates that global indices are used. This means that DE-local index
      space starts at the global lower bound for each DE.
\item [ESMC\_INDEX\_USER]
      Indicates that the DE-local index bounds are explicitly set by the user.
\end{description}

\subsection{ESMC\_LINETYPE}
\label{opt:lineType}

{\sf DESCRIPTION:\\}  This argument allows the user to select the path of the 
line which connects two points on the surface of a sphere.
This in turn controls the path along which distances are calculated and the 
shape of the edges that make up a cell.

The type of this flag is:

{\tt type(ESMC\_LineType\_Flag)}

The valid values are:
\begin{description}
\item [ESMC\_LINETYPE\_CART]
    Cartesian line. When this option is specified distances are calculated in a 
    straight line through the 3D Cartesian space in which the sphere is 
    embedded. Cells are approximated by 3D planes bounded by 3D Cartesian lines 
    between their corner vertices.
    When calculating regrid weights, this line type is currently the default for 
    the following regrid methods: ESMC\_REGRIDMETHOD\_BILINEAR,
    ESMC\_REGRIDMETHOD\_PATCH, ESMC\_REGRIDMETHOD\_NEAREST\_STOD, and 
    ESMC\_REGRIDMETHOD\_NEAREST\_DTOS.
\item [ESMC\_LINETYPE\_GREAT\_CIRCLE]
    Great circle line. When this option is specified distances are calculated 
    along a great circle path (the shortest distance between two points on a 
    sphere surface). Cells are bounded by great circle paths between their 
    corner vertices. When calculating regrid weights, this line type is 
    currently the default for the following regrid method: 
    ESMC\_REGRIDMETHOD\_CONSERVE.
\end{description}


\subsection{ESMC\_LOGKIND}
This flag is documented in section \ref{const:clogkindflag}.

\subsection{ESMC\_LOGMSG}
This flag is documented in section \ref{const:clogmsgflag}.

\subsection{ESMC\_MESHELEMTYPE}
This flag is documented in section \ref{const:cmeshelemtype}.

\subsection{ESMF\_METHOD}
\label{const:cmethod}

{\sf DESCRIPTION:\\}  
Specify standard ESMF Component method.

The type of this flag is:

{\tt type(ESMF\_Method\_Flag)}

The valid values are:
\begin{description}
\item [ESMF\_METHOD\_FINALIZE]
      Finalize method.
\item [ESMF\_METHOD\_INITIALIZE]
      Initialize method.
\item [ESMF\_METHOD\_READRESTART]
      ReadRestart method.
\item [ESMF\_METHOD\_RUN]
      Run method.
\item [ESMF\_METHOD\_WRITERESTART]
      WriteRestart method.
\end{description}

\subsection{ESMC\_POLEKIND}
This flag is documented in section \ref{const:cpolekind}.

\subsection{ESMC\_REDUCE}
\label{const:creduce}
{\sf DESCRIPTION:\\}
Indicates reduce operation.

The type of this flag is:

{\tt type(ESMC\_Reduce\_Flag)}

The valid values are:
\begin{description}
\item [ESMC\_REDUCE\_SUM]
      Use arithmetic sum to add all data elements.
\item [ESMC\_REDUCE\_MIN]
      Determine the minimum of all data elements.
\item [ESMC\_REDUCE\_MAX]
      Determine the maximum of all data elements.
\end{description}

\subsection{ESMC\_REGION}
\label{const:cregion}
{\sf DESCRIPTION:\\}
Specifies various regions in the data layout of an Array or Field object.

The type of this flag is:

{\tt type(ESMC\_Region\_Flag)}

The valid values are:
\begin{description}
\item [ESMC\_REGION\_TOTAL]
      Total allocated memory.
\item [ESMC\_REGION\_SELECT]
      Region of operation-specific elements.
\item [ESMC\_REGION\_EMPTY]
      The empty region contains no elements.
\end{description}

\subsection{ESMC\_REGRIDMETHOD}
This flag is documented in section \ref{opt:cregridmethod}.

\subsection{ESMC\_STAGGERLOC}
This flag is documented in section \ref{const:cstaggerloc}.

\subsection{ESMC\_TYPEKIND}
\label{const:ctypekind}

{\sf DESCRIPTION:\\}
Named constants used to indicate type and kind combinations supported by the
overloaded ESMC interfaces. The corresponding Fortran kind-parameter constants 
are described in the ESMF\_TYPEKIND section of Appendices of the ESMF Fortran 
reference manual.

The type of these named constants is:

{\tt type(ESMC\_TypeKind\_Flag)}

The named constants are:
\begin{description}
\item [ESMC\_TYPEKIND\_I1]
      Indicates 1 byte integer. \newline 
      (Only available if ESMF was built with 
      {\tt ESMF\_NO\_INTEGER\_1\_BYTE = FALSE}. This is {\em not} the default.)
\item [ESMC\_TYPEKIND\_I2]
      Indicates 2 byte integer. \newline
      (Only available if ESMF was built with 
      {\tt ESMF\_NO\_INTEGER\_2\_BYTE = FALSE}. This is {\em not} the default.)
\item [ESMC\_TYPEKIND\_I4]
      Indicates 4 byte integer.
\item [ESMC\_TYPEKIND\_I8]
      Indicates 8 byte integer.
\item [ESMC\_TYPEKIND\_R4]
      Indicates 4 byte real.
\item [ESMC\_TYPEKIND\_R8]
      Indicates 8 byte real.
\end{description}

\subsection{ESMC\_UNMAPPEDACTION}
\label{const:unmappedaction}
{\sf DESCRIPTION:\\}
Indicates what action to take with respect to unmapped destination points
and the entries of the sparse matrix that correspond to these points.

The type of this flag is:

{\tt type(ESMC\_UnmappedAction\_Flag)}

The valid values are:
\begin{description}
	\item[ESMC\_UNMAPPEDACTION\_ERROR]
	An error is issued when there exist destination points in a regridding
	operation that are not mapped by corresponding source points.
	\item[ESMC\_UNMAPPEDACTION\_IGNORE]
	Destination points which do not have corresponding source points are 
	ignored and zeros are used for the entries of the sparse matrix
	that is generated.
\end{description}

