\section{Conventions}
\label{sec:conventions}

\subsection{Document Conventions}

The following conventions for fonts and capitalization are used
in this document. \newline

\begin{tabular}{lll}
{\bf Style} & {\bf Meaning} & {\bf Example} \\ \hline
{\it italics}  & documents & {\it ESMF Reference Manual}\\
{\tt courier}  & code fragments & {\tt ESMF\_TRUE}\\
{\tt courier()}  & ESMF method name & {\tt ESMF\_FieldGet()} \\
{\bf boldface} & first definitions & An {\bf address space} is ...\\
{\bf boldface} & web links & {\bf Development} webpage \\
{Capitals}     & ESMF class name & DataMap \\
\end{tabular} 
 
ESMF class names frequently coincide with words commonly
used within the Earth system domain (field, grid, component, array, 
etc.)  The convention we adopt in this manual is that if a word is 
used in the context of an ESMF class name it is capitalized, and 
if the word is used in a more general context it remains in lower 
case.  We would write, for example, that an ESMF Field class 
represents a physical field.  

Diagrams are drawn using the Unified Modeling Language (UML).  UML 
is a visual tool that can illustrate the structure of 
classes, define relationships between classes, and describe sequences
of actions.  A reader interested in more detail can refer to a 
text such as {\it The Unified Modeling Language Reference Manual.}
 \cite{uml}

\subsection{Method Name and Argument Conventions}

There are conventions for how class methods are presented throughout 
this document.  Although Fortran interfaces are not case-sensitive,
we use case to help parse multi-word names.  We also use case to help
make the presentation of Fortran interfaces consistent with the 
presentation of C++ interfaces.  

Method names begin with {\tt ESMF\_}, followed by the class name, 
followed by the name of the operation being performed.  Each new 
word is capitalized. 

For method arguments that are multi-word, the first word is lower
case and subsequent words begin with upper case.  ESMF class 
names (including typed flags) are an exception.  When multi-word 
class names appear in argument lists, all letters after the first 
are lower case.  The first letter is lower case if the class is the
first word in the argument and upper case otherwise.  For 
example, in an argument list the DELayout class name may appear 
as {\tt delayout} or {\tt srcDelayout}.

Most Fortran calls in the ESMF are subroutines, with 
any returned values passed through the interface.  For the sake of 
convenience, some ESMF calls are written as functions.

A typical ESMF call thus looks like this:

\begin{verbatim}
call ESMF_<ClassName><Operation>(classname, firstArgument, 
           secondArgument, ..., rc)
\end{verbatim}

where \newline
{\tt <ClassName>} is the class name, \newline
{\tt <method>} is the name of the action to be performed, \newline 
{\tt classname} is a variable of the derived type associated 
with the class, \newline
the {\tt arg*} arguments are whatever other variables are required 
for the operation, \newline
and {\tt rc} is a return code. \newline

\subsection{Locating Methods in this Manual}

Methods for each class are located in the section devoted to 
that class in the {\it Reference Manual}.  In some classes, methods
are split into a number of different types.  For example, there are 
separate listings for Basic Field Methods, Field Overloads for Fortran Arrays,
and Field Communications.  The methods in each listing are ordered 
alphabetically.  The split into different listings is a side effect
of the automated document generation system we use; it reflects 
which methods are located in the same source files.  It is something 
we are working to eliminate!   


