% $Id: ESMF_usage.tex,v 1.7 2002/09/09 21:56:36 cdeluca Exp $

\subsection{Bindings}

The library includes both C/C++ and F90 bindings.  The C/C++ prefix
for procedures and parameters is {\tt ESMC\_} and the F90 prefix is
{\tt ESMF\_}.

\subsection{Object Function Conventions}

A large portion of the library is built using objects.  To standardize
the creation, use and destruction of these objects the library uses
certain common function for each object.

\begin{itemize}
\item{\bf ESM[F/C]\_<Class>New} Allocates a {\it deep} object.  This
  function will allocate space from the heap and may create other
  resources which must be free'd or deallocated.  All items created
  with {\bf New} must be deleted with {\bf Delete} below.  {\tt e.g.
    ESMF\_LogNew}
  
\item{\bf ESM[F/C]\_<Class>Delete} De-Allocates an object created with
  New.  Any transient resources that were created will be cleaned up.
  {\tt e.g. ESMF\_LogDelete}
  
\item{\bf ESM[F/C]\_<Class>Init} Initializes a {\it shallow} object.
  This class of objects does not need to be destructed and is
  guaranteed not to allocate any resources that must be cleaned up.
  {\tt e.g. ESMF\_TimeInit}
  
\item{\bf ESM[F/C]\_<Class>Set<Value>} Sets a given value with the
  class.  The {\bf Value} parameter is decided by the class.  {\tt
    e.g. ESMF\_LogSetState}
  
\item{\bf ESM[F/C]\_<Class>Get<Value>} Gets a given value with the
  class.  The {\bf Value} parameter is decided by the class.  {\tt
    e.g. ESMF\_LogGetState}
  
\item{\bf ESM[F/C]\_<Class>SetConfig} This function takes a list of
  resorces as defined in the resource section the class.  Some class
  may not have resources and this function has no meaning.  The
  function allows a user to set multiple resources with one function
  call.  {\tt e.g. ESMF\_LogSetConfig}
  
\item{\bf ESM[F/C]\_<Class>GetConfig} This function takes a list of
  resorces as defined in the resource section the class.  Some class
  may not have resources and this function has no meaning.  The
  function allows a user to get multiple resources with one function
  call.  {\tt ESMF\_LogGetConfig}
  
\item{\bf ESM[C]\_<Class>Construct} This function fills initializes an
  object with valid data.  This function is called by both the {\bf
    Init} and {\bf New} functions.  Depending on the type of object
  this function may or may not allocate resources that need to be
  freed.
  
\item{\bf ESM[C]\_<Class>Destruct} This function cleans up any
  resources that were created in the {\bf Construct} method.

\end{itemize}

\subsection{Constraints}

The library design imposes some constraints on the user:

\begin{itemize}
\item {\it No direct access of Fortran derived types.} Elements of
  Fortran derived types are private.
  
\item{\it All types should be initialized.} In order to provide
  consistent argument checking and to increase the overall robustness
  of the library, a user should call one of the {\tt ``Init''}
  routines before using the library's Fortran derived types, or the
  {\tt ``Construct''} routine before using its C/C++ classes.  A {\tt
    ``New''} routine is provided for C/C++ if dynamic memory
  allocation is desired.

\end{itemize}

\subsection{Error Handling}

All C/C++ procedures return an integer error code.  All F90 procedures have 
an optional integer return code argument (with the exception of a select few
functions that use {\tt stdargs}).  Return codes are translated 
into error descriptions using the methods: 

\begin{verbatim}
    void ESMC_ErrPrint(int rc)

    subroutine ESMF_ErrPrint(rc)  
    integer, intent(in), optional :: rc
\end{verbatim}

A return code of {\tt ESM[F/C]\_SUCCESS} indicates that an 
operation executed without errors.

The user can currently choose from two different error handlers.
{\tt ESM[F/C]\_ERR\_RETURN} will simply return from a routine in which an error 
is identified, without printing an error description.
{\tt ESM[F/C]\_ERR\_EXIT} will print a detailed error description including
file, function name, and line number, and will then terminate execution
(this is a simple exit, not an {\tt MPI\_ABORT}).  These handlers are set 
using the methods: 
\begin{verbatim}
    void _ErrHandlerSetType(ESM[F/C]_ErrHandlerType type)

    subroutine ESM[F/C]_ErrHandlerSetType(type)
    integer, intent(in) :: type
\end{verbatim}

The intent is to provide an error handling system in which
users can choose from a variety of handlers or supply their own.


\subsection{Attributes}

Attributes are a list of (name, value) pairs associated with any 
object in the library. The value can be of any data type. 
There are methods for listing, setting, and querying attribute lists.
These methods are expected to be implemented in the base class
and not specialized by higher level objects.








