\subsection{Demonstration Application}
\label{sec:demo}

In this section we describe the {\tt ESMF\_COUPLED\_WAVE} demonstration 
code that illustrates usage of both the ESMF infrastructure and superstructure. 
\subsubsection{Coupled Wave Demonstration Description}

The {\tt ESMF\_COUPLED\_WAVE} application is comprised of two ESMF 
{\tt Gridded Components}, defined on mutually exclusive processor sets, 
and a {\tt Coupler Component} defined on the union of those processor sets.  
The first Gridded Component, {\tt Solver}, solves the compressible 
time-dependent fluid flow equations.  The algorithm 
applies an explicit solution technique to a staggered, Arakawa C grid 
that is cartesian and uniform.  State variables, including density, 
pressure, viscosity and temperature, are located at cell-centers, while 
velocities are located at cell faces.  This component is initialized 
with a steady-state, one-dimensional flow.  The second Gridded Component, 
{\tt Injecter}, injects fluid into the first normal to the flow along 
one of the boundaries.  The injected fluid will have arbitrary velocity, 
temperature, density and duration, effectively setting some of 
the boundary conditions for the first component.  The {\tt Solver} and 
{\tt Injecter Components} sit on different cartesian grids.  The
{\tt Coupler Component} performs a regridding and transfers the 
boundary condition data from the {\tt Injecter} to the {\tt Solver}.

\subsubsection{Running the Coupled Wave Demonstration}

To run the demonstration, type 

\begin{verbatim}
gmake ESMF_COUPLED_FLOW
\end{verbatim}






























