% $Id$


The following compilers and utilities are required for compiling, linking and
testing the ESMF software. It is good common practice to use a consistent set
of Fortran/C++/C compilers from the same vendor, e.g. GNU, Intel, etc.
However, some vendor {\em combinations} of Fortran, C++, and C compilers,
e.g. Intel ifort with GNU g++, are also supported.
\begin{itemize}
\item Fortran compiler:
  \begin{itemize}
  \item GNU's gfortran v7.0 and newer, or
  \item Intel's ifort v18.0 and newer, or
  \item PGI's pgf90 v18.1 and newer, or
  \item NVHPC's nvfortran, or
  \item NAG's nagfor, or
  \item IBM's xlf, or
  \item CCE's ftn.
  \end{itemize}
\item C++ compiler:
  \begin{itemize}
  \item GNU's g++ v7.0 and newer, or
  \item Intel's icpc v18.0 and newer, or
  \item PGI's pgCC v18.1 and newer, or
  \item NVHPC's nvc++, or
  \item IBM's xlC, or
  \item CCE's CC.
  \item LLVM's clang
  \end{itemize}
\item C compiler:
  \begin{itemize}
  \item GNU's gcc v7.0 and newer, or
  \item Intel's icc v18.0 and newer, or
  \item PGI pgcc v18.1 and newer, or
  \item NVHPC's nvcc, or
  \item IBM's xlc, or
  \item CCE's cc.
  \item LLVM's clang
  \end{itemize}
\item MPI implementation compatible with the above compilers (but also see below
for the MPI-bypass build option):
  \begin{itemize}
  \item OpenMPI v3.0 and newer, or
  \item MPICH v2.1 and newer, or
  \item MVAPICH2 v2.0 and newer, or
  \item IntelMPI v18.0 and newer, or
  \item MPT 2.17 and newer, or
  \item CRAY-MPICH v7.7 and newer.
  \end{itemize}
\item GNU's \htmladdnormallink{gcc compiler}{http://gcc.gnu.org} -
for a standard cpp preprocessor implementation.
\item \htmladdnormallink{GNU Make}{http://www.gnu.org/software/make/make.html}.
\item \htmladdnormallink{Perl}{http://www.perl.com/download.csp} - for running
test scripts.
\end{itemize}

Internal packages that can optionally reference external libraries:
\begin{itemize}
\item LAPACK - version 3.x or newer
\item ParallelIO (PIO) - version 2.5.7 or newer
\item yaml-cpp - tag yaml-cpp-0.6.2 or newer
\end{itemize}

Optional external packages that must be specified for certain functions:
\begin{itemize}
\item NetCDF - version 3.6.x or newer (version 4.4 or newer required by PIO)
\item parallel-NetCDF - version 1.2.0 or newer (version 1.12 or newer required by PIO)
\item Xerces - version 3.1.0 or newer
\end{itemize}

ESMF can be built using a single-processor MPI-bypass library
that comes with ESMF by setting {\tt ESMF\_COMM=mpiuni}. This allows ESMF applications
to be linked and run in single-process mode.

In order to build html and pdf versions of the ESMF documentation, 
\htmladdnormallink{\LaTeX}{http://www.latex-project.org},
the \htmladdnormallink{latex2html}{http://www.latex2html.org}
conversion utility, and the Unix/Linux {\tt dvipdf} utility must be installed.
The csh shell is also required to complete the documentation build.
