% $Id$

\subsection{Running ESMF Examples}
\label{examples}


\subsubsection{Example source code}

Example source code for each class is found in the class's example
directory. For example, source code for the Time Manager class examples
are found in this directory:

\begin{verbatim}
        ESMF_DIR/src/Infrastructure/TimeMgr/examples/
\end{verbatim}

While the example code is formatted to be included in the documentation,
it also runs and compiles to ensure accuracy.  Examples generally 
contain simple usage of the basic methods for the class.

\subsubsection{Building and running examples}

The GNU makefile targets {\tt examples} and {\tt examples\_uni} build
and run programs found in a class's examples directory.  After the
examples are built, the {\tt examples} target runs the examples using
multiple processors, while {\tt examples\_uni} runs the examples on
a single processor.

These targets first build the ESMF library.

Run from ESMF\_DIR, this command will build and run all examples on
multiple processors:

\begin{verbatim}
       gmake examples
\end{verbatim}

If the command is run in an example source code directory, then only
the example from that directory will be built and run.  The examples
and output files are created in this directory:

\begin{verbatim}
ESMF_DIR/examples/examples$ESMF_BOPT/$ESMF_OS.$ESMF_COMPILER.$ESMF_ABI. \
	$ESMF_SITE/
\end{verbatim}

The name of an output file will begin with the name of the example
that created it followed by .stdout.

At the end of examples execution a script runs to analyze the results.

The script output indicates whether there are any example failures.
If any examples fail, please check if the failures are listed as known bugs in the ESMF release
page \htmladdnormallink{http://www.earthsystemmodeling.org/download/releases.shtml}{http://www.earthsystemmodeling.org/download/releases.shtml}
for your platform and compiler.
If the failures are not listed please contact ESMF Support at \htmladdnormallink{esmf\_support@list.woc.noaa.gov}
{mailto:esmf\_support@list.woc.noaa.gov}
Please indicate which examples are failing, and attach the output of the "gmake info" command to the email.


The following is a sample from the script output:

\begin{verbatim}



The following examples passed:


src/Infrastructure/Array/examples/ESMF_ArrayCreateEx.F90
src/Infrastructure/Array/examples/ESMF_ArrayGetEx.F90
src/Infrastructure/ArrayComm/examples/ESMF_ArrayCommEx.F90
src/Infrastructure/ArrayDataMap/examples/ESMF_ArrayDataMapEx.F90
src/Infrastructure/ArraySpec/examples/ESMF_ArraySpecEx.F90
src/Infrastructure/FieldBundle/examples/ESMF_FieldBundleCreateEx.F90
src/Infrastructure/FieldBundleDataMap/examples/ESMF_FieldBundleDataMapEx.F90
src/Infrastructure/DELayout/examples/ESMF_DELayoutEx.F90
src/Infrastructure/Field/examples/ESMF_FieldCreateEx.F90
src/Infrastructure/Field/examples/ESMF_FieldFromUserEx.F90
src/Infrastructure/Field/examples/ESMF_FieldGlobalEx.F90
src/Infrastructure/Field/examples/ESMF_FieldWriteEx.F90
src/Infrastructure/FieldComm/examples/ESMF_FieldCommEx.F90
src/Infrastructure/FieldDataMap/examples/ESMF_FieldDataMapEx.F90
src/Infrastructure/LogErr/examples/ESMF_LogErrEx.F90
src/Infrastructure/Regrid/examples/ESMF_RegridEx.F90
src/Infrastructure/Route/examples/ESMF_RouteEx.F90
src/Infrastructure/TimeMgr/examples/ESMF_AlarmEx.F90
src/Infrastructure/TimeMgr/examples/ESMF_CalendarEx.F90
src/Infrastructure/TimeMgr/examples/ESMF_ClockEx.F90
src/Infrastructure/TimeMgr/examples/ESMF_TimeEx.F90
src/Infrastructure/VM/examples/ESMF_VMAllFullReduceEx.F90
src/Infrastructure/VM/examples/ESMF_VMComponentEx.F90
src/Infrastructure/VM/examples/ESMF_VMDefaultBasicsEx.F90
src/Infrastructure/VM/examples/ESMF_VMGetMPICommunicatorEx.F90
src/Infrastructure/VM/examples/ESMF_VMScatterVMGatherEx.F90
src/Infrastructure/VM/examples/ESMF_VMSendVMRecvEx.F90
src/Superstructure/Component/examples/ESMF_AppMainEx.F90
src/Superstructure/Component/examples/ESMF_CplEx.F90
src/Superstructure/Component/examples/ESMF_GCompEx.F90
src/Superstructure/State/examples/ESMF_StateEx.F90
src/Superstructure/State/examples/ESMF_StateReconcileEx.F90


The following examples failed, did not build, or did not execute:


src/Infrastructure/Grid/examples/ESMF_GridCreateEx.F90
src/Infrastructure/TimeMgr/examples/ESMF_TimeIntervalEx.F90


The stdout files for the examples can be found at:
/home/bluedawn/svasquez/script_dirs/daily_builds/esmf/examples/
examplesO/AIX.default.64.default


Found 34 examples, 32 passed and 2 failed.


\end{verbatim}
