% $ Id: $

\documentclass[english]{article}
\usepackage{babel}
\usepackage{supertabular}
\usepackage{html}
\usepackage{times}
\usepackage{graphicx}
\usepackage[T1]{fontenc}

\newcommand{\docmttype}{Risk Report}
\newcommand{\req}[1]{\section{\hspace{.2in}#1}}
\newcommand{\sreq}[1]{\subsection{\hspace{.2in}#1}}
\newcommand{\ssreq}[1]{\subsubsection{\hspace{.2in}#1}}
\newcommand{\mytitle}{DRAFT \longname \docmttype ~~}

\newenvironment
{reqlist}
{\begin{list} {} {} \rm \item[]}
{\end{list}}

%===============================================================================
% User-defined commands
%-------------------------------------------------------------------------------
\newcommand{\longname}{ESMF RISKS }
\newcommand{\funcname}{RISKS }
\newcommand{\shortname}{RSK}
\newcommand{\myauthors}{ESMF Technical Leads:  V. Balaji, Cecelia DeLuca, Chris Hill}
%===============================================================================
\setlength{\textwidth}{6.5truein}
\setlength{\textheight}{8.5truein}
\setlength{\oddsidemargin}{0in}
\setlength{\unitlength}{1truecm}

\begin{document}

\bodytext{BGCOLOR=white LINK=#083194 VLINK=#21004A}

% Title page

\input{../../scripts/doc_templates/templates/title_alldoc}

\newpage
\tableofcontents

\newpage
\section{Synopsis}

This document lists a series of risks that the Earth System Modeling Framework project faces
in delivering a system that meets the requirements of a community standard with

\begin{enumerate}
\item adequate performance
\item general present and future applicability to a range of applications and a broad mix of hardware platforms
\item reasonable ease of adoption
\item inbuilt support for ongoing extensibility
\end{enumerate}

\subsection{Cataloging Risks}
Risks have been categorized as {\it Very High}, {\it High}, {\it Medium}, {\it Fairly Low} and {\it Low}.
This classification is used for both measuring impact on the ESMF project ( the {\it Level of Impact On Project} 
category) and for quantifying the currently perceived exposure to the risk within
the project ( the {\it Problem Likelihood} attribute), so that the
status of the risk is a pair of values, such as ({\it Low,Low}), that expresses
both the risk seriousness (the first element) and the risk impact (the second element) on
ESMF. A risk is recorded using the following template
\begin{itemize}
\item {\bf Short Title}: A brief, single sentence description of the risk.
\item {\bf Description}: Explanatory text that gives details of the risk.
\item {\bf Level of Impact On Project}: A single value that captures the impact on the project
if this risk occurs. Valid values are {\it Very High}, {\it High}, {\it Medium}, {\it Fairly Low} and
{\it Low}.
\item {\bf Nature of Impact}: Text that clarifies in what way the risk would impact the project.
\item {\bf Problem Likelihood}: A single value that captures the perceived probability that the risk
identified will occur. Valid values are {\it Very High}, {\it High}, {\it Medium}, {\it Fairly Low} and
{\it Low}.
\item {\bf Preventative Measures}: Text describing what measures are being used to deal with the
risk identified. The measures described under this heading are already in place.
\item {\bf Fix}: Text describing what actions have been identified as possible courses of action
to follow if the risk impact level or likelihood rise to high levels. These are courses of action 
that are not currently being pursued, but that could be adopted if the risk status warrants.
In some cases these actions could involve quite major shifts in project strategy.
\end{itemize}

\subsection{Risk Tracking and Review Procedures}
ESMF is funded on a strict budget and "hard" milestone basis. Special attention needs to be
paid to risks that affect the milestone schedule.
This document is reviewed on a monthly basis by the technical lead team and managerial adjustments
made as appropriate. Typical adjustments are expected to include updates to notes on specific risks,
occasional risk status adjustments and occasional addition of new risks.
Preventative measures are always to be taken for any risk that has a status greater than {\it Medium, Medium}.
Fix actions are taken for risks with a ({\it High, Medium}) status or higher, or where a risk
has a {\it Very High} status in either of the {\bf Level of Impact On Project} or the
{\bf Problem Likelihood} field.
The technical lead team will keep the NASA CT {\it milestone validation team} informed of 
risk issue status at their regular monthly telecons and will also involve project PI's where 
appropriate.

\subsection{Identifying Risks}
ESMF project members will be responsible for
\begin{itemize}
\item Proposing and assessing new risk issues.
\item Assessing progress in mitigating risks.
\item Reviewing and commenting on risk level.
\end{itemize}
These issues will be communicated via the standard 
project mechanisms of teleconferences and email to the technical lead team who will incorporate updates
into the document as appropriate.

\subsection{Document Organization}

The risks reported here are grouped into three parts. The first part {\bf ESMF Development Risks} describes
technical issues that have qualities that are tied to the nature of the planned ESMF development.
The second part {\bf ESMF Project Management Risks} deals with project management issues
that have qualities that are tied to the nature of the ESMF project. The final part, {\bf Generic Risks},
describes how standard project risks are being managed within ESMF.

\newpage

\part{ESMF Development Risks}

This part deals with risks that are related to the overall goal of the ESMF project and
to specific technical issues that ESMF faces.

\setcounter{section}{0}
\renewcommand{\thesection}{\shortname\arabic{section}}
\renewcommand{\thesubsection}{\shortname\arabic{section}.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\shortname\arabic{section}.\arabic{subsection}.\arabic{subsubsection}}

\req{Community Adoption Issues}
ESMF aims to be a widely used community standard. As such it faces a number of challenges that are
related to a mix of technical and organizational factors.
\sreq{Inadequate community dialog}
\begin{reqlist}
{\bf Level of Impact On Project:}  High \\
{\bf Nature of Impact:} ESMF is aiming to become a community standard. It won't be one if it isn't used.\\
{\bf Problem Likelihood:} Fairly Low\\
{\bf Preventive Measures Being Taken:} Community meetings. Active participation by project
members in related community efforts. Deployment of wide number of community codes
onto the framework. Attendance by project participants at related project
meetings. The Source Forge open source repository, along with its feedback forums
and a project web site with feedback forms also serve as conduits for community interaction.\\
{\bf Fix:} Increase community participation in project. Improve features to make
adoption more compelling.\\
{\bf Notes:} \\
\end{reqlist}

\sreq{Unable to respond to all community needs}
\begin{reqlist}
{\bf Level of Impact On Project:}  Medium \\
{\bf Nature of Impact:} If ESMF is perceived to be unresponsive to needs and unable to adapt to needs,
 community adoption will be in jeopardy.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Ongoing dialog to identify an prioritize
community needs. Critical requirements will be adopted to the extent
that contractual obligations admit. Track requirements that cannot be accommodated during the
project time frame. Plan and design for extensibility and future community contributions.\\
{\bf Fix:} Seek resources for requirements that cannot be accommodated within current
contract obligations. Discuss contract modifications with NASA if truly critical
need (as determined by advisory board members) can not be addressed. \\
{\bf Notes:} \\
\end{reqlist}

\sreq{Overcommitment to support community}
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Devoting resources to community support at the expense of contractual obligations
could impact milestone achievement.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Clear emphasis of project milestones. Implementation
team only accountable to technical leads.\\
{\bf Fix:} Down prioritize community needs that do not align with contractual obligations.
Discuss with NASA sponsors possible amendments to contract terms if appropriate.\\
{\bf Notes:} \\
\end{reqlist}

\sreq{Unable to achieve sufficient community adoption in project time frame}
The long-term viability of ESMF will rest on its broad use and a perception of
its value to a large user base. This will take time, possibly longer
than the three-year project window.
\begin{reqlist}
{\bf Level of Impact On Project:}  High \\
{\bf Nature of Impact:} Project could be perceived as having failed even though the
technical product it has created is effective and capable.\\
{\bf Problem Likelihood:} Fairly low \\
{\bf Preventative Measures Being Taken:} Broad range of dialog and visibility 
forums being maintained. Example codes, extensive documentation and tutorial
workshops are being planned. Involvement of broad range of groups in alpha and
beta testing. Adoption by Mechoso and Hauser external collaborators.
Close monitoring of experiences of collaborator groups.  
Working to develop follow-on funding sources for the project to continue
to add features requested by the community.
\\
{\bf Fix:} Solicit non-adopter input on what is missing or why framework is not working for them.
Review and revise design and implementation appropriately.\\
{\bf Notes:} \\
\end{reqlist}

\sreq{Framework too difficult and too complex for users.}
\begin{reqlist}
{\bf Level of Impact On Project:}  High \\
{\bf Nature of Impact:} Community adoption will not occur at sufficient level.\\
{\bf Problem Likelihood:} Fairly low \\
{\bf Preventative Measures Being Taken:}
Incremental development, review of prototype designs and feedback from users
will allow discovery of issues. Feedback will also be solicited through the
project web site. Independent integration by the Hauser and Mechoso investigator
teams and one independent adopter will provide additional validation.
The feedback will be used to focus the teams design and development efforts on 
specific parts of the framework that require clean, flexible and easy to use design.
\\
{\bf Fix:} Solicit non-adopter input on what is missing or why framework is not working for them.
Review and revise design and implementation appropriately.\\
{\bf Notes:} \\
\end{reqlist}

\req{Superstructure Level Issues}
\sreq{Arbitrary SPMD and MPMD support}
ESMF is required to manage and direct arbitrary collections of SPMD and MPMD programs 
running across a potentially heterogeneous mix of platforms.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Inability to achieve a sufficiently robust and maintainable
solution would limit range of code architectures that could work with ESMF.\\
{\bf Problem Likelihood:} Low \\
{\bf Preventative Measures Being Taken:} Detailed prototype studies to demonstrate all required
combinations.\\
{\bf Fix:} Reevaluate priorities after revisiting detailed requirements and
pare back initial functionality being implemented to only include
what is strictly required for milestones.\\
{\bf Notes:}
\end{reqlist}

\req{System Issues}
\sreq{Interoperation between languages}
ESMF will use a mix of Fortran, C and C++ languages. Several interlanguage binary binding
issues need to be addressed covering both normal execution and exceptional
behavior.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Multi-language support could create complex development
issues that draw resources away from other areas. It may also impose constraints
on possible designs and functionality that can be implemented.
\\
{\bf Problem Likelihood:} Low\\
{\bf Preventative Measures Being Taken:} Significant prototyping is being undertaken and
the language boundaries are being very clearly delineated.\\
{\bf Fix:} Entire framework would need to be in C/C++.\\
{\bf Notes:} Reimplementation in C/C++ would be time-consuming but
otherwise straightforward. However, complete C/C++ implementation could deter
some community members from adopting the framework.
\end{reqlist}

\sreq{F90 Continued Availability}
Fortran 90/95 remain {\it fringe} languages, even though most of the Earth system
community uses them. Two consequences of this are that ESMF Fortran 90 code will
rely on vendor supplied compilers that have an ever decreasing user base and
that there is no true open-source compiler for Fortran 90/95.
This impacts ESMF directly because some of its code will use F90 and
indirectly because many of its components exist in F90 only.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Code that is written in F90 would no longer be portable
and components would only be available on certain platforms. \\
{\bf Problem Likelihood:} Low\\
{\bf Preventative Measures Being Taken:} Some framework development is being pursued in C/C++.
Availability and quality of Fortran 90/95 compilers
is being tracked.\\
{\bf Fix:} Entire framework would need to be in C/C++.\\
{\bf Notes:} Reimplementation in C/C++ would be time-consuming but
otherwise straightforward. For components the impact would be far more
serious, but this is outside the scope of ESMF.
Components in Fortran 77, C and C++ would not be affected by lack of 
compiler availability as compilers for these languages exist on almost all hardware platforms.
\end{reqlist}

\sreq{Vector Platform Availability of ESMF}
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} The framework might not build and install "out of the box"
on a vector system.\\
{\bf Problem Likelihood:} Medium\\
{\bf Preventative Measures Being Taken:} ESMF is being tested and built on a range of *NIX systems. This will
ensure fairly easy portability to other systems, however, effort will still be needed 
to validate ESMF code and installation on a new platform. \\
{\bf Fix:} Test and modify to build on appropriate system\\
{\bf Notes:} This would require some resources but otherwise would not be a major challenge.
\end{reqlist}

\sreq{MS Windows Availability of ESMF}
\begin{reqlist}
{\bf Level of Impact On Project:} Fairly Low \\
{\bf Nature of Impact:} The framework might not build and install "out of the box"
on a Windows system.\\
{\bf Problem Likelihood:} High \\
{\bf Preventative Measures Being Taken:}  ESMF is being tested and built on a range of *NIX systems.
No specific tests of build actions are envisaged to ensure Windows compatibility.\\
{\bf Fix:} Test and modify to build on a Windows system\\
{\bf Notes:} This would require resources but otherwise would not be a major challenge.
\end{reqlist}

\sreq{Other OS Validation}
\begin{reqlist}
{\bf Level of Impact On Project:} Low \\
{\bf Nature of Impact:} The framework might not build and install "out of the box"
on a non-validated OS.\\
{\bf Problem Likelihood:} High \\
{\bf Preventative Measures Being Taken:} ESMF being tested and built on a range of *NIX systems\\
{\bf Fix:}  Test and modify to build on appropriate system\\
{\bf Notes:} This could require resources but otherwise would not be a major challenge.
However, for systems where no F90 compiler is available resource requirements could be substantial.
\end{reqlist}

\sreq{Insufficient test platform availability}
ESMF will need to be continually test on a range of platforms for
a variety of different hardware configurations and both small and larger
numbers of processors. This assumes that the project team will have easy
and continual access to a range of hardware.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Could impact milestone achievement by affecting code portability
and code performance.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:}
Resources have been requested from NASA and have recently been provided.
Usage, utility and responsiveness of the systems is being monitored.
\\
{\bf Fix:} Work to acquire special dedicated hardware resources for exclusive project
use.
Acquire compute resources through other channels - other agency centers,
industrial partners etc...
\\
{\bf Notes:} 
\end{reqlist}

\sreq{Performance Inconsistencies Between Platforms}
Multi-platform support introduces a risk of inconsistent performance across supported
platforms. System specific libraries and system calls, quality of compiler optimizations, and 
other items introduce machine specific performance problems. Vendor implementations
may not have optimal performance and it may be necessary to use a custom
implementation, an alternative set of libraries, etc... in order to 
achieve good performance on specific systems.
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Community members using a platform where the framework has
performance issues, will be less likely to adopt the framework.
and code performance.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:}
Performance profiling of the framework will be done
on each target platform to identify these issues early. Vendor contacts
are being established that will provide an effective route for 
system software problem reporting.
The framework is being designed to allow flexible adaptation to
different libraries so that alternate codes can be used with little
impact on end users.
\\
{\bf Fix:}For certain features the ESMF project could choose to develop its
own code or to incorporate extra, existing open-source code into the framework.
\\
{\bf Notes:} 
\end{reqlist}

\req{EVA code integration}
The EVA codes are being developed from many different legacy code sources.
Some of these codes are quite large and synthesizing the code into
a suite that is integrated into ESMF may involve considerable unanticipated work.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Could lead to schedule delays and milestone delays.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Experts from EVA code source institutions
are working with the core ESMF team EVA developers. A baseline EVA suite will
be assembled early in project.\\
{\bf Fix:} Increase involvement of source institutions in EVA suite
formulation. Arrange site visits to exchange expertise and to allow ESMF
team members to work with key developers at source institutions for relevant
codes. \\
{\bf Notes:} 
\end{reqlist}

\req{Infrastructure - Fields and Grid Level Issues}
\sreq{Too many complex features}
Many different grids will be supported under ESMF. A manageable and scalable approach
to development and testing is required to enable robust support for different grids.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Increased code complexity could lead to
difficulty of use, problems with portability and future evolution challenges.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} A layered interface has been defined 
which highlights generic grid properties and expresses them through common abstractions.
Many groups will attempt to deploy the system in their own in-house
applications and their experiences will be fed-back into the system design.\\
{\bf Fix:} Interface redesign \\
{\bf Notes:} 
\end{reqlist}

\sreq{Inadequate performance for certain grids}
Grid operations, like halo-update and transpose, can have significant impact on parallel scaling.
Regriddings between different grids can also greatly impact scaling.
A general purpose system that can scale and perform as well as specialized, application and
grid specific forms, could be hard to devise. Deficiencies in this arena will
impact adoption.
\begin{reqlist} 
{\bf Level of Impact On Project:} High \\ 
{\bf Nature of Impact:} Low performance would impact community adoption and
would impact milestone achievement.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Quantitative measures of grid primitive
performance are being devised and will be tracked.\\
{\bf Fix:} Develop specialized forms and increase level of resources devoted
to developing these parts of the system.\\
{\bf Notes:} 
\end{reqlist}

\sreq{Inadequate performance for certain platforms}
Grid operations can be specialized to take advantage of platform specific
architectural features. In a general purpose system it can be hard to compete with these
specialized forms. Low performance in this area could deter community adoption of ESMF.
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Low performance would impact community adoption and
would impact milestone achievement.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Quantitative measures of grid primitives
performance are being devised and will be tracked.\\
{\bf Fix:} Develop specialized forms and increase level of resources devoted
to developing these parts of the system. Reduce feature set for which
optimized support is developed.
\\
{\bf Notes:} 
\end{reqlist}

\req{Infrastructure - Utilities Level Issues}
Inadequate resources for developing highly portable and versatile
signaling, performance monitoring, I/O, communication, memory sharing and namespace sharing tools.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} This would affect the robustness, portability and usability
of ESMF. \\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Tools developed under other projects (for example PetSC, ALICE, TAU)
will be adopted wherever possible.
\\
{\bf Fix:} Devote more resources to this area. Reduce feature set for which
highly optimized support is provided initially. \\
{\bf Notes:} Dependencies of ESMF on third-party software installation will be noted
in the Software Engineering Plan system documentation.
\end{reqlist}

\newpage
\part{ESMF Project Management Risks}
This part deals with risks that are related to the overall goal of the ESMF project and
to specific project management issues that arise from the goals and project structure.


\req{Distributed Project Team}
ESMF involves many institutions and a geographical distributed team. 
This can create competing time pressures and difficulty in communication.

\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Communication difficulties could result in system quality and feature deficiencies. \\
{\bf Problem Likelihood:} Low \\
{\bf Preventative Measures Being Taken:} Detailed scheduling. Frequent (bi-weekly) telephone
conferencing. Quarterly meetings. Location of core framework implementation and development
team in one place. Easy sharing of managed development code through source forge. 
Extensive use of e-mail lists. Frequent progress reviews and open dialog on project status.
\\
{\bf Fix:} Hosting of team members at NCAR where needed.
Site visits by development team members will be undertaken as needed.
\\
{\bf Notes:} 
\end{reqlist}

\sreq{Overambitious scheduling of new development}
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Slippages in scheduling could delay dependent development and deployment efforts and could impact milestone payments.  \\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Early prototyping efforts to explore most
complex areas of development.
\\
{\bf Fix:} Reductions in feature set if needed.
\\
{\bf Notes:} 
\end{reqlist}


\sreq{Coding Standards may be ignored}
The core framework and EVA suite are required to meet very specific agreed code quality and coding
and documentation standards for milestone acceptance.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:}  Failure to develop code that meets the
standard will cause project funding delays.
 \\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:} Developers are aware of the
coding standards, as described in the Software Developers Guide, and of the peer review and 
independent {\it milestone validation} processes that have been established.
\\
{\bf Fix:} Closer involvement of {\it milestone validation team} in ongoing code 
assessment. Development of automated tools to assist in code standard validation.
\\
{\bf Notes:} 
\end{reqlist}


\sreq{Cross-project milestone coordination}
Several phases in the project rely on earlier milestones under a separate
project.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Dependent phases could be seriously impacted by earlier delays.
\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:} Joint technical lead team has representation
from all projects. Weekly ESMF wide telecons are used to track upcoming
milestones. Project phases are planned jointly and time lines mapped out
together at joint meetings. Technical lead team tracks milestone
work-schedule and coordinates multiple milestone countdown reviews prior 
to milestones.
\\
{\bf Fix:} Reassign milestone work responsibilities as appropriate.
Increase NASA CT {\it milestone validation team} involvement in pre-milestone
countdown reviews.
\\
{\bf Notes:} 
\end{reqlist}

\newpage
\part{Generic Project Risks}

This part documents how more general risk issues, that are faced by all complex projects,
are addressed with ESMF.

\req{Staff skill issues}
Task estimates may be based on assumptions about individual skills that become unavailable
due to unforeseen conditions.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Replacement staff have to be hired and trained which affects
project schedule.\\
{\bf Problem Likelihood:} Medium \\
{\bf Preventative Measures Being Taken:} Ensure that multiple developers are familiar
with each system element. Ensure documentation is kept up to date and coding
standards are adhered to.\\
{\bf Fix:} \\
{\bf Notes:} 
\end{reqlist}

\req{Schedule delays with cascade effects}
A delay in one task may cause cascading delays in other dependent tasks.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Could impact milestone target dates.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:} For each project phase
(milestones E, F, G and K) a schedule is developed at a whole project meeting. 
Schedules are recorded in MS project and monitored by the technical lead team.
Week by week management is carried out by the technical lead team with reference to the
agreed schedule. This management involves detailed scheduling of
work reviews among sub-groups. This reviews are carried out via telephone
with Source Forge used to share material where distributed team members are
involved.  \\
{\bf Fix:} Introduce more aggressive scheduling strategies, including
increased site traveling by technical leads.\\
{\bf Notes:} 
\end{reqlist}

\req{Code quality}
Inadequate adherence to encapsulation and modularity principles could
lead to overly fragile software.
\begin{reqlist}
{\bf Level of Impact On Project:} High \\
{\bf Nature of Impact:} Could impact milestone achievement. System reliability, framework ease of use
and maintainability can all be effected.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:} Peer review and testing of all modules.
\\
{\bf Fix:}May require redesign and reimplementation.\\
{\bf Notes:} 
\end{reqlist}

\req{System stability}
As code evolves previously working feature may break.
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Could impact milestone achievement.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:} Automated regression test system will catch 
broken code at a fine grain level. Peer review of modules and adherence to modular
design principles will also help catch issues here.
\\
{\bf Fix:}\\
{\bf Notes:} 
\end{reqlist}

\req{Inappropriate design}
Inappropriate design may require redesign and re-implementation.
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Could impact milestone achievement.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:}
Prototyping to identify and explore potential designs prior to committing to
an implementation.
\\
{\bf Fix:}\\
{\bf Notes:} 
\end{reqlist}

\req{Team conflicts}
Conflicts between team members may result in poor communication, poor designs,
and extra rework.
\begin{reqlist}
{\bf Level of Impact On Project:} Fairly Low \\
{\bf Nature of Impact:} Could impact milestone achievement.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:}
Identify any conflicts between team members early in the project and work to
resolve the issues. Maintain an open, participatory process.
\\
{\bf Fix:}\\
{\bf Notes:} 
\end{reqlist}

\req{Lack of critical skills}
Personnel with critical skills may not be available due to funding, other
project commitments, etc...
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Could impact milestone achievement.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:}
Identify personnel with critical skills and ensure other team
members are familiar with their domain of expertise.
Ensure that personnel with critical skills are committed
to the development effort and confirm availability.
New hires are being selected on the basis of skill sets that are
missing and can be trained, both formally and informally, as needed.
\\
{\bf Fix:}\\
{\bf Notes:} 
\end{reqlist}

\req{Inaccurate Progress Tracking}
Inaccurate progress tracking may result in not knowing the project is
behind schedule until late in the project.
\begin{reqlist}
{\bf Level of Impact On Project:} Medium \\
{\bf Nature of Impact:} Could impact milestone achievement.\\
{\bf Problem Likelihood:} Fairly Low \\
{\bf Preventative Measures Being Taken:}
Weekly project and technical lead teleconferences
include a discussion of project time line.
\\
{\bf Fix:}\\
{\bf Notes:} 
\end{reqlist}

\end{document}
