
\section{Release Notes}

The purpose of ESMF v1.0 is to provide a first look at the ESMF
Application Programming Interface (API), and to demonstrate the viability 
of the ESMF architecture and implementation.  For this release, we have focused 
on implementing major architectural features and basic functionality.  The 
current performance characteristics and memory requirements of the software 
are unlikely to resemble those in later releases.

Our next major release, ESMF v2.0, will occur in April 2004.  At that time, 
the {\it ESMF User Guide} will be expanded to include a comprehensive section 
on how to adapt application codes for the framework, and support staff will be 
available to assist users with ESMF adoption.  We do not anticipate preparing other public
releases or patches for v1.0 between now and the next release - our focus
will be on developing an easy to use, production-quality product.  We are 
relying on Earth system modelers to provide us with the feedback necessary 
to realize this goal.  Section \ref{sec:Support} of this document includes 
instructions on submitting comments on ESMF v1.0 to our development team.

\section{What is the Earth System Modeling Framework?}

The Earth System Modeling Framework (ESMF) is a structured collection of 
software building blocks that can be used or customized to develop 
Earth system model components, and assemble them into applications.  
The simplest view of the ESMF is that it consists of an
{\it infrastructure} of utilities and data structures for creating 
model components, and a {\it superstructure} for coupling them.  
User code sits between these two layers, making calls to the infrastructure
libraries beneath it and being scheduled and synchronized by the 
superstructure above it.  The configuration resembles a sandwich, as
shown in Figure 1.

The ESMF architecture is scalable, flexible paradigm for building highly 
complex climate, weather, and related applications from components such
as atmospheric models, land models, and data assimilation systems.  The 
ESMF is not a single master application into which all components must fit; 
rather it is a way of developing components so that they can be used 
in many different user-written applications.  Model components that adopt 
ESMF are usable in different contexts without code modification, and may be
incorporated into other ESMF-based modeling systems within the Earth 
science community.  In addition to high-level organization, ESMF provides 
a set of robust, portable, performance optimized libraries for regridding, 
data transfers, I/O, time management, and other common modeling functions.  
ESMF users may choose to extensively rewrite their codes to take advantage 
of the ESMF infrastructure, or they may decide to simply wrap user-written 
components in ESMF interfaces in order to adopt the ESMF architecture and 
utilize framework coupling services.

\section{The ESMF Reference Manual for Fortran90}

This {\it ESMF Reference Manual} is a listing of ESMF standard interfaces
for Fortran90.  Interfaces are grouped by class.  A class is an 
object-oriented software design construct that embodies a specific 
concept like a physical field.  (For more on object-oriented
design, how we have implemented object-oriented design in Fortran, and 
how it is used to organize 
ESMF, please see the \htmladdnormallink{{\it ESMF Architecture Document}}
{http://www.esmf.ucar.edu/esmf_docs/ESMF_archdoc}.

The {\it Reference Manual} is divided into two main parts.  The 
first part is a description 
of the classes in the ESMF superstructure, and the second part is a 
description of classes in the ESMF infrastructure.  The major class
presented in the superstructure is the Component class, which can 
be of type Gridded Component, Coupler Component, or Application Component.


For a broad overview of how the ESMF software works, see the 
\htmladdnormallink{{\it ESMF User Guide}}{http://www.esmf.ucar.edu}.














