% $Id: ESMF_desc.tex,v 1.4 2003/04/04 15:11:55 cdeluca Exp $

\section{Description}

The ESMF is a structured collection of software building blocks 
that can be used or customized to develop model components, assemble them into an 
application, and run the application.  Unlike a software library, in a framework some
software elements are partly defined or unimplemented.  For example, a framework
may require that a gridded model provide a {\tt GetGrid} method, and specify 
a format for describing the grid, but leave the implementation of the method up
to the framework user.  This relatively non-intrusive sort of standardization enables 
different model components to interact with each other, and also
use general methods provided by the framework (for example a framework supplied
{\tt Regrid} method).

The functional split between developing and combining components is the
foremost feature of the ESMF architecture.  The simplest view of the ESMF 
is that it consists of a layered {\it infrastructure} of utilities and data 
structures for building model components and a {\it superstructure} for coupling 
and running them.  User-provided components fit between these 
layers like the filling in a sandwich.  Each successively higher layer builds
on elements of the one below it.  Utilities such as communication
primitives, logging and profiling tools are used in methods that
manipulate field and grid data structures.  These data structures may be
tightly integrated into user-supplied components, or they may simply 
help to wrap a component's native data structures in order to provide 
a standard interface for coupling.  The superstructure provides a systematic 
approach to sequencing and synchronizing user-supplied components that 
interact to form an application.  The main layers in the ESMF architecture are 
shown in Figure~\ref{fig:sandwich}.  

%\subsection{Shared Objects}
\input{ESMF_shareobj}



































