% $Id: Attribute_rest.tex,v 1.21 2011/01/05 20:05:41 svasquez Exp $
%
% Earth System Modeling Framework
% Copyright 2002-2011, University Corporation for Atmospheric Research,
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics
% Laboratory, University of Michigan, National Centers for Environmental
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory,
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.


\subsubsection{Attributes}

\begin{itemize}
\item Case insensitive Attribute names, conventions, purposes, and values will be enabled in a future release.
\end{itemize}


\subsubsection{Attribute packages}

\begin{itemize}
\item A future capability may be to automatically create default object Attribute packages upon ESMF object creation.
\item The implemention of Grids is still in flux within the CIM.  This will effect the final appearance of the Gridspec package in both ESMF and ESG. It is anticipated that an additional CIM Grid Attribute Package will be created.
\item A CIM Scientific Property Attribute Package will be added. For CMIP5, hundreds of Scientific Properties have been identified. All of these will be added to ESMF. 
\item The Attribute packages ISO Responsible Party, ISO Citation, and CIM Platform can only be created automatically within a CIM Main component Attribute package.  In a future release, it will be possible to create these within other CIM Attribute packages as required, or as separate, standalone packages.
\end{itemize}


\subsubsection{Attribute hierarchies}

\begin{itemize}
\item The option of "deep" copies of an Attribute hierarchy will be added.
\end{itemize}


\subsubsection{Attribute import and export}
\begin{itemize}
\item The CIM XML output in this release validates against CIM development v1.5, svn revision 2210, which was current as of this ESMF release's freeze date, 9/19/2010.  CIM v1.5 development is continuing, with an official v1.5 release expected within weeks after this ESMF release.  ESMF, in a future release, will conform to the officially released CIM v1.5 or later.
\item The CIM XML output format, as described above, is ingestable into ESG also as of this ESMF release's freeze date, 9/19/2010. However, ESG, like CIM, is in flux with active development.  When ESG officially releases, future releases of ESMF will be compatible with it.
\item Only the CIM XML format is ingestable by ESG, not ESMF XML nor tab-delimited.
\item CIM Attribute packages can only be output (to CIM XML); they may be inputtable (via XML) in a future release.
\item The ESMF grid Attribute package XML output file only contains the nested GridSpec Attributes; Attributes RegDecompX and RegDecompY will be added in a future release. 
\end{itemize}
