% $Id$
%
% Earth System Modeling Framework
% Copyright 2002-2022, University Corporation for Atmospheric Research, 
% Massachusetts Institute of Technology, Geophysical Fluid Dynamics 
% Laboratory, University of Michigan, National Centers for Environmental 
% Prediction, Los Alamos National Laboratory, Argonne National Laboratory, 
% NASA Goddard Space Flight Center.
% Licensed under the University of Illinois-NCSA License.


The ESMF Geom class is used as a container for other ESMF geometry objects (e.g. an ESMF Grid). This allows a generic
representation of a geometry to be passed around (e.g. through a coupled system) without it's specific type being known.
Some operations are supported on a Geom object and more will be added over time as needed. However, if
an unsupported operation is needed, then the specific geometry object can always be pulled out and operated on that way.

In addition to the geometry object, a Geom can also contain information describing a location on a geometry. For example, in the case of
a Grid, a geometry object will also contain a stagger location. Having this location information allows the creation of Fields and
other capabilities to be performed in the most generic way on a Geom object. For geometries where it is appropriate, the user can
optionally specify this location information during the creation of a Geom object. However, if no location is specified, then default values for
this information are provided which match those which would be used when creating a Field with the specific geometry
(e.g. stagger location ESMF\_STAGGERLOC\_CENTER for a Grid).
