% $Id$

%\subsection{Usage and Examples}

\subsubsection{Fortran unit number management} \label{fio:unitnumbers}
The {\tt ESMF\_UtilIOUnitGet()} method is provided so that applications
using ESMF can remain free of unit number conflicts --- both when combined
with other third party code, or with ESMF itself.  This call is typically
used just prior to an {\tt OPEN} statement:

\begin{verbatim}
  call ESMF_UtilIOUnitGet (unit=grid_unit, rc=rc)
  open (unit=grid_unit, file='grid_data.dat', status='old', action='read')
\end{verbatim}

By default, unit numbers between 50 and 99 are scanned to find an unopened
unit number.

Internally, ESMF also uses {\tt ESMF\_UtilIOUnitGet()} when it needs to open
Fortran unit numbers for file I/O.  By using the same API for both user and
ESMF code, unit number collisions can be avoided.

When integrating ESMF into an application where there are conflicts with
other uses of the same unit number range, such as when hard-coded unit number
values are used, an alternative unit number range can be specified.
The {\tt ESMF\_Initialize()} optional arguments {\tt IOUnitLower} and {\tt IOUnitUpper}
may be set as needed.  Note that {\tt IOUnitUpper} must be set to a value higher than
{\tt IOUnitLower}, and that both must be non-negative.  Otherwise {\tt ESMF\_Initialize}
will return a return code of {\tt ESMF\_FAILURE}.  ESMF itself does not typically need more
than about five units for internal use.

\begin{verbatim}
  call ESMF_Initialize (..., IOUnitLower=120, IOUnitUpper=140)
\end{verbatim}

All current Fortran environments have preconnected unit numbers, such as
units 5 and 6 for standard input and output, in the single digit range.
So it is recommended that the unit number range is chosen to begin at unit 10
or higher to avoid these preconnected units.

\subsubsection{Flushing output}

Fortran run-time libraries generally use buffering techniques to improve I/O
performance.  However output buffering can be problematic when output is needed,
but is ``trapped'' in the buffer because it is not full.
This is a common occurance when debugging a program, and inserting {\tt WRITE} statements
to track down the bad area of code.  If the program crashes before the output
buffer has been flushed, the desired debugging output may never be seen --- giving
a misleading indication of where the problem occurred.  It would be desirable
to ensure that the output buffer is flushed at predictable
points in the program in order to get the needed results.
Likewise, in parallel code, predictable flushing of output buffers is a common
requirement, often in conjunction with {\tt ESMF\_VMBarrier()} calls.  

The {\tt ESMF\_UtilIOUnitFlush()} API is provided to flush a unit as desired.  Here is
an example of code which prints debug values, and serializes the output to a
terminal in PET order:

\begin{verbatim}
  type(ESMF_VM) :: vm

  integer :: tty_unit
  integer :: me, npets

  call ESMF_Initialize (vm=vm, rc=rc)
  call ESMF_VMGet (vm, localPet=me, petCount=npes)

  call ESMF_UtilIOUnitGet (unit=tty_unit)
  open (unit=tty_unit, file='/dev/tty', status='old', action='write')
  ...
  call ESMF_VMBarrier (vm=vm)
  do, i=0, npets-1
    if (i == me) then
      write (tty_unit, *) 'PET: ', i, ', values are: ', a, b, c
      call ESMF_UtilIOUnitFlush (unit=tty_unit)
    end if
    call ESMF_VMBarrier (vm=vm)
  end do
\end{verbatim}
