% $Id: Grid_rest.tex,v 1.6 2004/06/22 20:54:17 jwolfe Exp $

%\subsubsection{Restrictions and Future Work}

\begin{enumerate}

\item {\bf Support is limited to 3D, logically rectangular grids.} 
Currently the only interfaces supported are for three-dimensional, 
logically rectangular grids. 

\item {\bf Support is limited to 2D grid distributions.} The 
decomposition of grids is limited to two dimensions.

\item {\bf Future Grid Create methods.}  Currently grids can only
be created by internal generation.  In the future, the following
create methods will be added:
 \begin{description}
    \item [ESMF\_GridCreateCopy]
          Create a new Grid by copying another Grid
    \item [ESMF\_GridCreateCutout]
          Create a new Grid as a subset of an existing Grid
    \item [ESMF\_GridCreateDiffRes]
          Create a new Grid by coarsening or refining an existing Grid
    \item [ESMF\_GridCreateExchange]
          Create a new Grid from the intersection of two existing grids
    \item [ESMF\_GridCreateRead]
          Create a new Grid by reading in from a file
 \end{description}

\item {\bf Future Grid types.}  The following grids will
be supported, although only some will have internal generation 
methods:
 \begin{description}
    \item [ESMF\_GRID\_TYPE\_LATLON\_GAUSS]
          Latitude/Longitude grid with gaussian-spaced latitudes
    \item [ESMF\_GRID\_TYPE\_LATLON\_MERC]
          Latitude/Longitude grid with Mercator-spaced latitudes
    \item [ESMF\_GRID\_TYPE\_REDUCED]
          Latitude/Longitude grid where the number of longitudinal points is a
          function of the latitude
    \item [ESMF\_GRID\_TYPE\_DIPOLE]
          Displaced-pole dipole grid
    \item [ESMF\_GRID\_TYPE\_TRIPOLE]
          Tripolar grids
    \item [ESMF\_GRID\_TYPE\_DATASTREAM]
          Data stream - set of locations
    \item [ESMF\_GRID\_TYPE\_PHYSFOURIER]
          Mixed Fourier Space/Physcal Space grid
    \item [ESMF\_GRID\_TYPE\_SPHER\_SPECT]
          Spectral space for spherical harmonics
    \item [ESMF\_GRID\_TYPE\_CART\_SPECT]
          Spectral space for cartesian coordinates
    \item [ESMF\_GRID\_TYPE\_GEODESIC]
          Spherical geodesic grid
    \item [ESMF\_GRID\_TYPE\_CUBEDSPHERE]
          Cubed sphere grid
    \item [ESMF\_GRID\_TYPE\_EXCHANGE]
          Intersection of two grids, which is itself a grid
 \end{description}

\item {\bf Future coordinate system support.}  Support for the following
coordinate systems will be added:
 \begin{description}
    \item [ESMF\_COORD\_SYSTEM\_USER]
          User-defined coordinate system
    \item [ESMF\_COORD\_SYSTEM\_CYLINDRICAL]
          Cylindrical coordinates
    \item [ESMF\_COORD\_SYSTEM\_LATFOURIER]
          Mixed latitude/Fourier spectral space
    \item [ESMF\_COORD\_SYSTEM\_SPECTRAL]
          Wavenumber space
    \item [ESMF\_COORD\_SYSTEM\_LAGRANGIAN]
          Lagrangian coordinates
 \end{description}

\item {\bf Future horizontal Grid staggerings.}  Support for the following
horizontal staggerings will be added:
 \begin{description}
    \item [ESMF\_GRID\_HORZ\_STAGGER\_E]
          Arakawa E
    \item [ESMF\_GRID\_HORZ\_STAGGER\_Z]
          C grid equivalent for geodesic grid
 \end{description}

\item {\bf Future vertical Grid types.}  Support for the following
vertical grids will be added:
 \begin{description}
    \item [ESMF\_COORD\_SYSTEM\_DEPTH]
          Vertical z coordinate depth (0 at top surface)
    \item [ESMF\_COORD\_SYSTEM\_PRESSURE]
          Vertical pressure coordinate
    \item [ESMF\_COORD\_SYSTEM\_SIGMA]
          Vertical sigma coordinate
    \item [ESMF\_COORD\_SYSTEM\_THETA]
          Vertical theta coordinate
    \item [ESMF\_COORD\_SYSTEM\_ETA]
          Vertical eta coordinate
    \item [ESMF\_COORD\_SYSTEM\_ISOPYCNAL]
          Vertical density coordinate
    \item [ESMF\_COORD\_SYSTEM\_HYBRID]
          Hybrid vertical coordinates
    \item [ESMF\_COORD\_SYSTEM\_LAGRANGIAN]
          Lagrangian coordinates
 \end{description}

\item {\bf Future Grid masks.}  Grid masks will be implemented, including
support for the following mask types:
 \begin{description}
    \item [ESMF\_GRID\_MASKTYPE\_LOGICAL]
          Logical mask
    \item [ESMF\_GRID\_MASKTYPE\_MULT]
          Multiplicative mask
    \item [ESMF\_GRID\_MASKTYPE\_REGION\_ID]
          Integer assigning unique ID to each point
 \end{description}

\end{enumerate}


