% $Id: Grid_options.tex,v 1.4 2004/03/23 05:00:11 cdeluca Exp $

\subsubsection{ESMF\_GridType}

{\sf DESCRIPTION:\\}
Grid types supported by ESMF.

Valid values are:
\begin{description}
   \item [ESMF\_GridType\_LatLon]
         Equally-spaced lat/lon grid.
   \item [ESMF\_GridType\_Unknown]
         Unknown or undefined grid.

%   \item [ESMF\_GridType\_Mercator] Mercator lat/lon grid
%   \item [ESMF\_GridType\_Dipole] Displaced-pole dipole grid
%   \item [ESMF\_GridType\_Tripole] Tripolar grids
%   \item [ESMF\_GridType\_XY] Cartesian equally-space x-y grid
%   \item [ESMF\_GridType\_DataStream] Data stream
%   \item [ESMF\_GridType\_PhysFourier] Mixed Fourier Space/Phys Space grid
%   \item [ESMF\_GridType\_LatLonGauss] lat/lon grid with Gaussian latitudes
%   \item [ESMF\_GridType\_Reduced] reduced grid
%   \item [ESMF\_GridType\_SphericalSpectral] spectral space for spherical harmonics
%   \item [ESMF\_GridType\_Geodesic] spherical geodesic grid
%   \item [ESMF\_GridType\_CubedSphere] cubed sphere grid
\end{description}

\subsubsection{ESMF\_GridStagger}

{\sf DESCRIPTION:\\}
Grid staggerings supported by ESMF.

Valid values are:
\begin{description}
   \item [ESMF\_GridStagger\_A]
         Arakawa A (centered velocity).
   \item [ESMF\_GridStagger\_B]
         Arakawa B (velocities at grid corner).
   \item [ESMF\_GridStagger\_C]
         Arakawa C (velocities at cell faces).
   \item [ESMF\_GridStagger\_D]
         Arakawa D.
   \item [ESMF\_GridStagger\_E]
         Arakawa E.
   \item [ESMF\_GridStagger\_Z]
         C grid equivalent for geodesic grid.
   \item [ESMF\_GridStagger\_VertCenter] 
         Vertical velocity at vertical midpoints.
   \item [ESMF\_GridStagger\_VertFace]
         Vertical velocity/Pgrad at top(bottom)face.
   \item [ESMF\_GridStagger\_Unknown]
         Unknown or undefined staggering.

\end{description}

\subsubsection{ESMF\_CoordOrder}

{\sf DESCRIPTION:\\}
The Grid class can be set to different default ordering of coordinates, for
example KIJ.  This parameter describes ordering options supported by ESMF.

Valid values are:
\begin{description}
   \item [ESMF\_CoordOrder\_XYZ]
         IJK maps to XYZ.
   \item [ESMF\_CoordOrder\_XZY]
         IJK maps to XZY.
   \item [ESMF\_CoordOrder\_YXZ]
         IJK maps to YXZ.
   \item [ESMF\_CoordOrder\_YZX]
         IJK maps to YZX.
   \item [ESMF\_CoordOrder\_ZXY]
         IJK maps to ZXY.
   \item [ESMF\_CoordOrder\_ZYX]
         IJK maps to ZYX.
   \item [ESMF\_CoordOrder\_Unknown]
         Unknown or undefined coordinate ordering.
\end{description}

\subsubsection{ESMF\_CoordSystem}

{\sf DESCRIPTION:\\}
Supported coordinate systems.

Valid values are:
\begin{description}
   \item [ESMF\_CoordSystem\_User]
         User-defined coordinate system.
   \item [ESMF\_CoordSystem\_Spherical]
         Spherical coordinates (lon,lat).
   \item [ESMF\_CoordSystem\_Cartesian]
         Cartesian coordinates (x,y).
   \item [ESMF\_CoordSystem\_Cylindrical] 
         Cylindrical coordinates.
   \item [ESMF\_CoordSystem\_LatFourier]
         Mixed latitude/Fourier spectral space.
   \item [ESMF\_CoordSystem\_Spectral]
         Wavenumber space.
   \item [ESMF\_CoordSystem\_Depth]
         Vertical z coordinate depth (0 at top surface).
   \item [ESMF\_CoordSystem\_Height]
         Vertical z coordinate height (0 at bottom).
   \item [ESMF\_CoordSystem\_Pressure]
         Vertical pressure coordinate.
   \item [ESMF\_CoordSystem\_Sigma] 
         Vertical sigma coordinate.
   \item [ESMF\_CoordSystem\_Theta]
         Vertical theta coordinate.
   \item [ESMF\_CoordSystem\_Eta]
         Vertical eta coordinate.
   \item [ESMF\_CoordSystem\_Isopycnal]
         Vertical density coordinate.
   \item [ESMF\_CoordSystem\_Hybrid]
         Hybrid vertical coordinates.
   \item [ESMF\_CoordSystem\_Lagrangian] 
         Lagrangian coordinates.
   \item [ESMF\_CoordSystem\_Unknown]
         Unknown or undefined coordinate system.
\end{description}

\subsubsection{ESMF\_GridMaskType}

{\sf DESCRIPTION:\\}
Supported types of grid masks.

Valid values are:
\begin{description}
   \item [ESMF\_GridMaskType\_Logical]
         Logical mask.
   \item [ESMF\_GridMaskType\_Mult]
         Multiplicative mask.
   \item [ESMF\_GridMaskType\_RegionID]
         Integer assigning unique ID to each point.
   \item [ESMF\_GridMaskType\_Unknown]
         Unknown or undefined mask type.
\end{description}
